### 2016-01-04 code/Rmd/MSSM_RNASeq.Rmd
Minor changes in MSSB RNASeq covariates code

---
title: "Covariates analysis of MSBB RNASeq data"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./MSSM_RNASeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB RNASeq')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

### Clear R console screen output
cat("\014")

## Load required libraries
library(synapseClient)
library(rGithubClient) ## Needs the dev branch
library(ggplot2)
library(reshape2)
library(data.table)
library(plyr)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(limma)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB RNASeq samples';

thisFileName <- 'MSSM_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Data were obtained from AMP-AD portal of MSSM, and following brain regions were analysed for differential expression between case and control:
  | Region | Name                      | 
  | :----- | ------------------------- | 
  | BM_10  | Frontal Pole (FP)         |
  | BM_22  | Superior Temporal Gyrus   |
  | BM_36  | Parahipocampal Gyrus (PHG)|

* BM 10: frontopolar prefrontal cortex, rostrolateral prefrontal cortex, or anterior prefrontal cortex, 
* BM 22: In humans it corresponds approximately to the lateral and caudal two thirds of the superior temporal gyrus, 
* BM 36: Ectorhinal area 36 is a subdivision of the cytoarchitecturally defined temporal region of cerebral cortex.

Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn3205353'
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID)
COUNT <- read.table(COUNT_OBJ@filePath,header=T,sep='\t',check.names = F, row.names = 1)

# Get metadata
METADATA_ID <- 'syn3205337'
METADATA_OBJ <- synGet(METADATA_ID)
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA <- read.table(METADATA_OBJ@filePath,sep='\t',header=T, row.names=1)
METADATA <- cbind(METADATA, data.frame(Dx = ifelse(METADATA$CDR<1,'Control','AD')))
```
### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

Remove genes that have less than 1 cpm counts in at least 50% of samples.
```{r cpmnormalisation, cache=TRUE}
ind <- intersect(rownames(METADATA), colnames(COUNT))
METADATA <- METADATA[ind,]
COUNT <- COUNT[, ind]
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT,MIN_GENE_CPM=1, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
```
### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c('TISSUE','Donor','RACE','CDR','CERJ','SEX','NP.1','bbscore','Dx')
ContCovariates <- c('PMI','AOD','PlaqueMean')

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- rownames(METADATA)

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)

COVARIATES <- dplyr::mutate(COVARIATES, TISSUE.Dx = paste(TISSUE,Dx,sep='.'))
COVARIATES$TISSUE.Dx <- factor(COVARIATES$TISSUE.Dx)
rownames(COVARIATES) <- rownames(METADATA)
```
Covariate correlation
```{r covariates.correlation, fig.width=10, fig.height=10}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 1)
ggheatmap.show(COVARIATES.CORRELATION$plot, col.width=0.3, row.width=0.15)
```
### Normalisation (with NULL)
Initial normalisation usign voom (with NULL design)
```{r initial.voom.normalisation}
# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(PROCESSED_COUNTS$filteredExprMatrix, design=NULL, plot=T)

# Initial design matrix with covariates under consideration
DM = getDesignMatrix(COVARIATES, Intercept = F)
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
DESIGN = getDesignMat2Fact(DM$design,FactorCovariates)
rownames(DESIGN) = rownames(METADATA)
preAdjustedSigCovars = runPCAandPlotCorrelations(VOOM.GENE_EXPRESSION$E, DESIGN,'NULL design(voom-normalized)', isKeyPlot=TRUE)

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL, fig.width=15, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[1]]$plotData
```
### Normalisation (custom/manual design)
Re-perform normalisation with custom design matrix (manually with "TISSUE.Dx","SEX","RACE","PMI" as fixed effects and Donor as random effects)
```{r reperform.normalisation, cache=TRUE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,c("TISSUE.Dx","SEX","RACE","PMI"),drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

# Estimate correlation of random effects
VOOM.ADJUSTED.GENE_EXPRESSION = voom(PROCESSED_COUNTS$filteredExprMatrix, design=DM1$design, plot=F)
correlation <- duplicateCorrelation(VOOM.ADJUSTED.GENE_EXPRESSION$E, design=DM1$design, block=COVARIATES$Donor)

# Re-calculate voom weights with correlation of random effects
VOOM.ADJUSTED.GENE_EXPRESSION = voom(PROCESSED_COUNTS$filteredExprMatrix, design=DM1$design, plot=F,
                                     block=COVARIATES$Donor, correlation = correlation$cor)

# Re-calculate linear fit using new weights and new design
VOOM.ADJUSTED.FIT = lmFit(VOOM.ADJUSTED.GENE_EXPRESSION$E, design=DM1$design, weights = VOOM.ADJUSTED.GENE_EXPRESSION$weights,
                          block=COVARIATES$Donor, correlation = correlation$cor)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,VOOM.ADJUSTED.GENE_EXPRESSION$E)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
rownames(DESIGN) <- rownames(METADATA)
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION,
                                              dplyr::select(DESIGN,-starts_with('Donor')),
                                              'all adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
```

```{r residualSigCovars.manual, fig.width=15, fig.height=8}
residualSigCovars[["PC_res"]][[1]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "TISSUE.Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB RNASeq", parentId = parentId)
CODE <- synStore(CODE)

# Expression
expr = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsembleIDs')
write.table(expr, file='expression.tsv', sep = '\t', row.names=F, quote=F)
obj = File('expression.tsv', 
           name = 'Voom Adjusted Weighted Residual Expression (Tissue.Dx added)', 
           parentId = CODE$properties$id)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)

# Covariates
covar = rownameToFirstColumn(COVARIATES, 'SampleId')
write.table(covar, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
obj = File('covariates.tsv', 
           name = 'Covariates', 
           parentId = CODE$properties$id)
obj = synStore(obj,
               used = ALL_USED_IDs, 
               activityName = activityName, 
               executed = thisFile, 
               activityDescription = activityDescription)
```