### 2016-01-07 code/Rmd/ROSMAP_RNASeq.Rmd
Added markdown to process RNASeq files from ROSMAP cohorts from AMPAD knowledge portal

---
title: "Covariate analysis of ROSMAP RNASeq"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./ROSMAP_RNASeq.Rmd",
                                 parentId ="syn5570291",
                                 entityName = 'ROSMAP RNASeq')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

## Needs the dev branch
library(rGithubClient)

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570291';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of ROSMAP RNASeq data';

thisFileName <- 'ROSMAP_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn3505720';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID)
COUNT <- read.table(COUNT_OBJ@filePath,header=T,sep='\t',check.names = F)

# COUNT_ID1 <- 'syn3505720'; #'syn3505732'
# ALL_USED_IDs <- COUNT_ID1
# COUNT_OBJ1 <- synGet(COUNT_ID1)
# COUNT1 <- read.table(COUNT_OBJ1@filePath,header=T,sep='\t',check.names = F)
# 
# # Download expression data
# COUNT_ID2 <- 'syn3505724'
# ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_ID2
# COUNT_OBJ2 <- synGet(COUNT_ID2)
# COUNT2 <- read.table(COUNT_OBJ2@filePath,header=T,sep='\t',check.names = F)

# Get clinical metadata
METADATA.CLINICAL_ID <- 'syn3191087'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID
METADATA.CLINICAL_OBJ <- synGet(METADATA.CLINICAL_ID)
METADATA.CLINICAL <- read.table(METADATA.CLINICAL_OBJ@filePath,sep=',',header=T)

# Get technical covariates
# METADATA.TECH <- read.csv('./ROSMAP_RNA_Seq_RIN_Sample_READS.csv')
# METADATA.TECH <- METADATA.TECH[!duplicated(METADATA.TECH$Projid),]
METADATA.TECH_ID <- 'syn4300313'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.TECH_ID
METADATA.TECH_OBJ <- synGet(METADATA.TECH_ID)
METADATA.TECH <- read.table(METADATA.TECH_OBJ@filePath,sep='\t',header=T)

# Match technical and clinical covariates
METADATA <- dplyr::inner_join(METADATA.CLINICAL,METADATA.TECH)
```

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  mutate(Sampleid = paste(Sampleid, Batch, sep = '_')) %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi))  
```

```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$Sampleid, colnames(EXPR))

EXPR <- COUNT; #merge(COUNT1,COUNT2,by = 'gene_id',all=T)
rownames(EXPR) <- EXPR$gene_id
EXPR <- EXPR[, colnames(EXPR) %in% indToRetain]

# Match covariates to expression data
rownames(METADATA) = METADATA$Sampleid
METADATA = METADATA[indToRetain,]
```
Dorsolateral prefrontal cortex of `r dim(EXPR)[2]` subjects from the ROS and MAP cohorts were obtained from AMP-AD portal

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
#"braaksc","ceradsc","cts_mmse30_first_ad_dx","cts_mmse30_lv"
FactorCovariates <- c("Batch","msex","apoe_genotype", "race", "spanish", "cogdx")
ContCovariates <- c("RINcontinuous","age_first_ad_dx","age_death","age_at_visit_max","pmi","educ")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Covariate correlation
```{r covariates.correlation, fig.width=10, fig.height=10}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.1)
ggheatmap.show(COVARIATES.CORRELATION$plot, col.width=0.3, row.width=0.15)
```

### Gene Filtering
Retain genes with atleast 50% of the samples have >= 0 FPKM counts
```{r gene.filtering, cache=TRUE}
# Remove genes
EXPR <- EXPR[(rowSums(EXPR>0)/dim(EXPR)[2] >= 0.5),]

# To avoid taking log of zero add 1e-4 to the FPKM counts
EXPR <- log2(EXPR+1e-3)
```
`r dim(EXPR)[1]` number of genes and `r dim(EXPR)[2]` number of samples are considered for the analysis

### Significant Covariates
Correlation between pca of normalised mRNA expression and covariates
```{r preAdjusted.covariates, cache=TRUE, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
DM = getDesignMatrix(COVARIATES, Intercept = F)
DESIGN = getDesignMat2Fact(DM$design,FactorCovariates)
rownames(DESIGN) = rownames(COVARIATES)
preAdjustedSigCovars = runPCAandPlotCorrelations(EXPR, DESIGN,'non-normalized', isKeyPlot=TRUE)

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=10, fig.height=6}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom design)
Fitting a linear model with cogdx, Batch, RINcontinuous, and pmi 
```{r iterative.normalisation, cache=TRUE, results='asis', echo=FALSE}
postAdjustCovars <- c('cogdx','Batch',"RINcontinuous","pmi")

writeLines(paste('Using following covariates in the final model',paste(postAdjustCovars,collapse=','),'as fixed effects',sep=' '))

# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

# Re-calculate linear fit using new weights and new design
FIT = lmFit(EXPR, design=DM1$design)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(FIT,EXPR)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
rownames(DESIGN) <- rownames(COVARIATES)
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION,
                                              DESIGN,
                                              'all adjusted design',
                                              isKeyPlot=TRUE)
```

```{r residualSigCovars.manual.ALL, fig.width=15, fig.height=8}
residualSigCovars[["PC_res"]][[2]]$plotData
```

```{r addVarsBack}
varsToAddBackIn <- c("cogdx1",  "cogdx2", "cogdx3", "cogdx4", "cogdx5", "cogdx6")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + (FIT$coefficients[,varsToAddBackIn] %*% t(data.matrix(DESIGN[,varsToAddBackIn])))
RESIDUAL.GENE_EXPRESSION = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsemblGeneID') %>%
  tidyr::separate(col = EnsemblGeneID, into = c('ensembl_gene_id','position'), sep = '\\.')
```

```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = data.frame(cogdx = 1:6, No.OfSamples = summary(COVARIATES$cogdx))
kable(tmp)
```

Interpretation of cogdx scores

1.  NCI, No cognitive impairment (No impaired domains)
2.  MCI, Mild cognitive impairment (One impaired domain) and NO other cause of CI
3.  MCI, Mild cognitive impairment (One impaired domain) AND another cause of CI
4.  AD, Alzheimer's disease and NO other cause of CI (NINCDS PROB AD)
5.  AD, Alzheimer's disease AND another cause of CI (NINCDS POSS AD)
6.  Other dementia. Other primary cause of dementia

```{r get.gene.symbols, cache=TRUE}
# Define biomart object
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- Folder(name = "Differential Expression Analysis ROSMAP RNASeq (Based on cogdx)", parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'Sampleid_batch')
write.table(COVARIATES, file = 'covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store covariates
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'FPKM Normalised Adjusted Residual Expression (cogdx added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXPRP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |