### 2016-01-25 code/Rmd/Mayo_Cerebellum_Array.Rmd
Added markdown for processing cerebellum microarray data from mayo-ufl-isb

---
title: "Covariate analysis of UFL-Mayo-ISB cerebellum arrays"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./Mayo_Cerebellum_Array.Rmd",
                                 parentId ="syn5570325",
                                 entityName = 'UFL_Mayo_ISB Cerebellum Microarray')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

## Needs the dev branch
library(rGithubClient)

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB Cerebellum array data';

thisFileName <- 'Mayo_Cerebellum_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
EXPR_ID <- 'syn3256501';
ALL_USED_IDs <- EXPR_ID
EXPR_OBJ <- synGet(EXPR_ID)
EXPR <- fread(EXPR_OBJ@filePath,sep=',',data.table=F)

# Get clinical metadata
METADATA_ID <- 'syn3256502'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA_OBJ <- synGet(METADATA_ID)
METADATA <- read.table(METADATA_OBJ@filePath,sep=',',header=T, stringsAsFactors=F)
```

### Data preprocessing
```{r preprocess.data,cache=TRUE}
rownames(EXPR) <- EXPR$IID
EXPR <- data.matrix(EXPR[,-(1:2)]) %>% t

# Fix metadata
rownames(METADATA) <-  METADATA$IID
METADATA = METADATA[colnames(EXPR),]

METADATA = dplyr::mutate(METADATA,
                         plate0 = 6*plate0,
                         plate1 = 1*plate1,
                         plate2 = 2*plate2,
                         plate3 = 3*plate3,
                         plate4 = 4*plate4,
                         plate = plate0+plate1+plate2+plate3+plate4)
METADATA$plate[METADATA$plate == 0] = 5
METADATA$plate[METADATA$plate == 6] = 0
```
Cerebellum of `r dim(EXPR)[2]` subjects from the Mayo cohorts were obtained from [AMP-AD portal](`r EXPR_ID`)

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c("Dxn", "Sex",  "E4dose", "plate")
ContCovariates <- c("Age", "RIN", "RINsqAdj")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$IID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Covariate correlation
```{r covariates.correlation, fig.width=10, fig.height=10}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.1)
ggheatmap.show(COVARIATES.CORRELATION$plot, col.width=0.3, row.width=0.15)
```

### Outlier analysis
Clustering of raw data
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(EXPR, scale.=T)

# Plot first 2 PCs
plotdata <- data.frame(IID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, METADATA, by="IID")

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Dxn), shape=as.factor(plate), size=as.numeric(Age)))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= IID), size=4, hjust=0)
p

COVARIATES.tmp = data.matrix(COVARIATES)
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
#                     dendroLabels = FALSE,
                    abHeight = 10, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
DM = getDesignMatrix(COVARIATES, Intercept=F)
DESIGN = getDesignMat2Fact(DM$design,FactorCovariates)
colnames(DESIGN) = substr(colnames(DESIGN),1,50)
preAdjustedSigCovars = runPCAandPlotCorrelations(EXPR, DESIGN,
                                                 'NULL design',
                                                 isKeyPlot=TRUE)

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),
                                     preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r paste(gsub('_','\\\\_',adjustCovars), collapse= ',')`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=25, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (iterative)
Since many covariates are correlated, re-normalising COVARIATES with an iterative design matrix.
```{r iterative.adjusted.voom.normalisation, cache=TRUE, results='asis', echo=FALSE}
postAdjustCovars = c('Dxn')

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars

significantCovars = designMatVarsToCovars(getCovariatesMap(DM), 
                                          residualSigCovars$significantCovars)

covariatesEffects = designMat2CovEffects(getCovariatesMap(DM),
                                         residualSigCovars$Effects.significantCovars)

covariatesEffects = covariatesEffects[setdiff(significantCovars, postAdjustCovars)]

postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))

loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 100){
  writeLines(paste('Using following covariates in the model:', 
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Fit linear model
  FIT = lmFit(EXPR, 
              design=DM1$design)

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(FIT,EXPR)

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), postAdjustCovars)

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                DESIGN,
                                                'all adjusted design',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  significantCovars = designMatVarsToCovars(getCovariatesMap(DM), residualSigCovars$significantCovars)
  covariatesEffects = designMat2CovEffects(getCovariatesMap(DM),residualSigCovars$Effects.significantCovars)
  covariatesEffects = covariatesEffects[setdiff(significantCovars,postAdjustCovars)]

  postAdjustCovars = c(postAdjustCovars,names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <-paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=','), 'as fixed effects')

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

```{r adjusted.expression, results='asis', echo=FALSE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F], Intercept = F)
tmp = linColumnFinder(DM1$design)

writeLines('Degenerate variables are:')
writeLines(tmp$relations)
DM1$design = DM1$design[,tmp$indepCols]

# Linear model with new design
ADJUSTED.FIT = lmFit(EXPR, design=DM1$design)  
```
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r calculate.residuals, include=FALSE}
# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT,
                                              EXPR)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              DESIGN,
                                              'residual matrix of all adjusted design',
                                              isKeyPlot=TRUE)
```
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
```{r addVarsBack}
varsToAddBackIn <- c("Dxn0", "Dxn1")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  ADJUSTED.FIT$coefficients[,varsToAddBackIn] %*% t(data.matrix(DM1$design[,varsToAddBackIn]))
RESIDUAL.GENE_EXPRESSION = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'illumina_humanht_12_v4') 
```

```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$Dxn))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
Where 0 is Non-Ad and 1 is AD
```{r get.gene.symbols, cache=TRUE}
# Define biomart object
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Query biomart
Illumina2HGNC <- getBM(attributes = c("illumina_humanht_12_v4", "hgnc_symbol"),
                       filters = "illumina_humanht_12_v4", 
                       values = RESIDUAL.GENE_EXPRESSION$illumina_humanht_12_v4,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(Illumina2HGNC, RESIDUAL.GENE_EXPRESSION) %>%
  dplyr::select(illumina_humanht_12_v4, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum Microarray', parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'ID')
write.table(COVARIATES, file = 'covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store expression values
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'Voom Adjusted Residual Expression (Dxn added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |