### 2016-02-29 code/Rmd/synNetGen.Rmd
Added special ACL privileges to gold standard

---
title: "Synthetic network and data generation using GeneNetWeaver(GNW)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./synNetGen.Rmd",
                                 parentId ="syn5569099",
                                 entityName = 'Synthetic Networks')
```
#### Clone GNW repo from github
```{r getGNW, eval=FALSE}
wd = getwd()
setwd('../../../')

# Clone GNW repo
system('git clone https://github.com/tschaffter/gnw.git')

setwd(wd)
```
#### Initialise libraries
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(synapseClient)
library(data.table)
library(dplyr)
library(stringr)
library(knitr)

## Needs the dev branch
library(githubr)

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5569099';
activityName = 'Synthetic network generation';
activityDescription = 'Synthetic network data from ecoli using GNW';

thisFileName <- 'synNetGen.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='synNet')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))

# Create folder to store results
FOLD = Folder(name = 'Synthetic Networks', parentId = parentId)
FOLD = synStore(FOLD)
```
#### Extract network
```{r extractNet}
# Extract settings text file
con = file('../../../gnw//sandbox/settings.txt')
settings.txt = readLines(con)
close(con)

netSize = c(a = 1e1, b = 1e2, c = 1e3, d = 1e4)
for (net in names(netSize))){
  # Number of minimum regulators
  minReg = max(5, round(netSize[[net]]*0.5))

  # Create output directory
  outputDir = paste0('SubNetSize',net)
  system(paste0('mkdir ', outputDir))

  # Write settings file to the output directory
  con = file(paste0(outputDir,'/settings.txt'))
  writeLines(settings.txt, con)
  close(con)

  # Extract 10 subnetworks of size n from ecoli data
  system(paste('java','-jar','../../../gnw/gnw-3.1.2b.jar',
                 '--extract','-c',paste0(outputDir,'/settings.txt'), '--input-net',
                 '../../../gnw/sandbox/ecoli_transcriptional_network_regulonDB_6_7.tsv',
                 '--random-seed','--greedy-selection','--num-subnets=10',
                 paste0('--subnet-size=',netSize[[net]]),'--output-net-format=4',
                 paste0('--output-path=',outputDir),
                 paste0('--min-num-regulators=',minReg)))

  # Convert generated subnetworks to tsv data
  FNames = dir(outputDir, pattern='*.xml',full.names=T)
  for (fnames in FNames){
    system(paste('java','-jar','../../../gnw/gnw-3.1.2b.jar',
                 '--transform','-c',paste0(outputDir,'/settings.txt'), '--input-net',
                 fnames, '--output-net-format=0',
                 paste0('--output-path=',outputDir)))
  }

  # Change settings for simulations  
  tmp = settings.txt
  tmp[grep('outputDirectory = ', tmp)] = paste('outputDirectory =',outputDir)
  tmp[grep('ssDualKnockouts',tmp)] = "ssDualKnockouts = 0"
  tmp[grep('tsDREAM4TimeSeries',tmp)] = "tsDREAM4TimeSeries = 0"

  con = file(paste0(outputDir,'/settings.txt'))
  writeLines(tmp, con)
  close(con)

  # Simulate from extracted networks
  for (fnames in FNames){
    system(paste('java','-jar','../../../gnw/gnw-3.1.2b.jar',
                 '--simulate','-c',paste0(outputDir,'/settings.txt'), 
                 '--input-net', fnames, '--output-path',outputDir))

    # Read all data into one file
    allFiles = dir(pattern = 'ecoli')
    system(paste('mv',allFiles[grep('.xml',allFiles)],outputDir))
    system(paste('mv',allFiles[grep('goldstandard.tsv',allFiles)],outputDir))

    fileName = tools::file_path_sans_ext(fnames) %>% str_replace(paste0(outputDir,'/'),'')

    data = list()

    extractData <- function(pat,allfiles, file.name){
      read.table(allfiles[grep(paste0(file.name,pat),allfiles)], header=T, sep = '\t', quote='')
    }

    allPatterns1 = c('_wildtype.tsv','_knockdowns.tsv', '_multifactorial.tsv', '_multifactorial_perturbations.tsv', '_knockouts.tsv')
    data[['1']] = lapply(allPatterns1, extractData, allFiles, fileName) %>%
      do.call(rbind,.)

    allPatterns2 = c('_nonoise_wildtype.tsv','_nonoise_knockdowns.tsv', '_nonoise_multifactorial.tsv', '_nonoise_knockouts.tsv')
    data[['2']] = lapply(allPatterns2, extractData, allFiles, fileName) %>%
      do.call(rbind, .)

    allPatterns3 = c('_noexpnoise_wildtype.tsv','_noexpnoise_knockdowns.tsv', '_noexpnoise_multifactorial.tsv', '_noexpnoise_knockouts.tsv')
    data[['3']] = lapply(allPatterns3, extractData, allFiles, fileName) %>%
      do.call(rbind, .)

    # Store data in synaspe
    fileNum = str_split(str_split(fnames, '-')[[1]][2],'\\.')[[1]][1]
    SUB.FOLD = Folder(name = paste0(outputDir, '_', fileNum), parentId = FOLD$properties$id)
    SUB.FOLD = synStore(SUB.FOLD)

    objs = lapply(names(data), function(x, Data, subFolder){
      tmp = data.matrix(Data[[x]]) %>% t %>% rownameToFirstColumn('GeneNames')
      colnames(tmp)[-1] = paste0('EXP',colnames(tmp)[-1])

      # Store data in synapse
      write.table(tmp, file = paste0('Expression_',x,'.tsv'), sep = '\t', row.names=F, quote=F)
      OBJ = File(paste0('Expression_',x,'.tsv'), 
                 name = paste('Expression',x), 
                 parentId = subFolder$properties$id)
      OBJ = synStore(OBJ, activityName = activityName,
                     executed = thisFile, activityDescription = activityDescription)
    }, data, SUB.FOLD)

    OBJ = File(paste0(tools::file_path_sans_ext(fnames), '.tsv'),
               name = 'Gold Standard', parentId = SUB.FOLD$properties$id)
    OBJ = synStore(OBJ, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

    # Set private acl to the gold standard file
    resAcc = ResourceAccess(principalId = 3319864, 
                            accessType = c('CREATE', 'READ', 'UPDATE', 'DELETE', 'CHANGE_PERMISSIONS', 'DOWNLOAD', 'UPLOAD', 'PARTICIPATE', 'SUBMIT', 'READ_PRIVATE_SUBMISSION', 'UPDATE_SUBMISSION', 'DELETE_SUBMISSION', 'TEAM_MEMBERSHIP_UPDATE', 'SEND_MESSAGE', 'CHANGE_SETTINGS'))
    acl = AccessControlList(id = OBJ$properties$id, resourceAccess = ResourceAccessList(resAcc))
    tmp = synCreateEntityACL(acl)

    system('rm ecoli*')
  }
}
```