### 2016-06-11 code/Rmd/PredictPMI.Rmd
Added markdown to identify features which can predict PMI in ROSMAP study

---
title: "Prediction model for PMI using ROSMAP Reprocessed RNASeq"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./PredictPMI.Rmd",
                                 parentId ="syn5570291",
                                 entityName = 'Predict PMI')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(Biobase)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570291';
activityName = 'Predict PMI';
activityDescription = 'Predict PMI using ROSMAP data';

thisFileName <- 'PredictPMI.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r rosmap.download.data, cache=TRUE}
# Download reprocessed counts (paired end)
COUNT_PE_ID <- 'syn6112813';
ALL_USED_IDs <- COUNT_PE_ID
COUNT_PE <- read.table(synGet(COUNT_PE_ID)@filePath, header=T, sep='\t', check.names = F)

# Download reprocessed counts (single end)
COUNT_SE_ID <- 'syn6117775';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_SE_ID
COUNT_SE <- read.table(synGet(COUNT_SE_ID)@filePath, header=T, sep='\t', check.names = F)

COUNT = COUNT_PE + COUNT_SE[rownames(COUNT_PE), colnames(COUNT_PE)]
rownames(COUNT) = sapply(rownames(COUNT), function(x){unlist(str_split(x, '\\.'))[1]})

# Download filtered logCPM
LOG.CPM_ID <- 'syn6129370';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_ID
LOG.CPM <- read.table(synGet(LOG.CPM_ID)@filePath, header=T, sep='\t', check.names = F)

COUNT = COUNT[intersect(rownames(COUNT), LOG.CPM$ensembl_gene_id), intersect(colnames(COUNT), colnames(LOG.CPM))]

# Get covariates
COVARIATES_ID <- 'syn6129368'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = COVARIATES_ID
COVARIATES <- read.table(synGet(COVARIATES_ID)@filePath,sep='\t',header=T, row.names = 1)

FactorCovariates <- c("cogdx")
ContCovariates <- c("RINcontinuous","age_death","pmi")

COVARIATES = COVARIATES[,c(FactorCovariates,ContCovariates),drop=F]
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates,drop=F], factor)

# Get surrogate variables
SURR.VAR_ID <- 'syn6140323'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = SURR.VAR_ID
SURR.VAR <- read.table(synGet(SURR.VAR_ID)@filePath,sep='\t',header=T)
rownames(SURR.VAR) = rownames(COVARIATES)
COVARIATES = cbind(COVARIATES, SURR.VAR)

COUNT = COUNT[,rownames(COVARIATES)]
```
Using dorsolateral prefrontal cortex of `r dim(COUNT)[2]` subjects from the ROS and MAP cohorts to predict PMI. Following sample are removed `r paste(removedIDs, collapse = ',')`
### Fit model with PMI
```{r w.pmi}
writeLines(paste('Using following covariates in the model:',paste(colnames(COVARIATES), collapse=', '),'as fixed effects'))
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES,Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

# Estimate sample weights using voom methodology
VOOM.ADJUSTED.GENE_EXPRESSION = voom(COUNT, design=DM1$design, plot=F)

# Fit linear model using estimated weights and calculate BIC's
BIC.WPMI = sapply(1:dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1], function(gene){
  y = VOOM.ADJUSTED.GENE_EXPRESSION$E[gene,]
  w = VOOM.ADJUSTED.GENE_EXPRESSION$weights[gene,]
  x = DM1$design
  obj = lm(y~0+., data = cbind(x,data.frame(y=y)), weights = w)
  BIC(obj)
})
names(BIC.WPMI) = rownames(VOOM.ADJUSTED.GENE_EXPRESSION$E)
```
### Fit model with out PMI
```{r wo.pmi}
COVARIATES.tmp = COVARIATES[,-(4)]
writeLines(paste('Using following covariates in the model:',paste(colnames(COVARIATES.tmp), collapse=', '),'as fixed effects'))

# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES.tmp,Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

# Estimate sample weights using voom methodology
VOOM.ADJUSTED.GENE_EXPRESSION = voom(COUNT, design=DM1$design, plot=F)

# Fit a linear model
VOOM.ADJUSTED.FIT = lmFit(VOOM.ADJUSTED.GENE_EXPRESSION)

# Calculate residual
RESIDUAL.EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT, VOOM.ADJUSTED.GENE_EXPRESSION$E)

# Fit linear model using estimated weights and calculate BIC's
BIC.WOPMI = sapply(1:dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1], function(gene){
  y = VOOM.ADJUSTED.GENE_EXPRESSION$E[gene,]
  w = VOOM.ADJUSTED.GENE_EXPRESSION$weights[gene,]
  x = DM1$design
  obj = lm(y~0+., data = cbind(x,data.frame(y=y)), weights = w)
  BIC(obj)
})
names(BIC.WOPMI) = rownames(VOOM.ADJUSTED.GENE_EXPRESSION$E)
```
### Predict PMI
```{rpredict.pmi}
# Extract features that show improvement in the model with pmi when compared to without pmi
genes = rownames(VOOM.ADJUSTED.GENE_EXPRESSION$E)[which((BIC.WOPMI-BIC.WPMI) >= 6)]

# Extract x and y
x = t(RESIDUAL.EXPRESSION[genes,])
y = COVARIATES$pmi

# Remove samples with 0 pmi
ind = y == 0
x = x[!ind,]
y = log(y[!ind])

# Linear regression model
glm.model = foreach(i = 1:100,
                .export = c('x','y'),
                .packages = c('glmnet', 'CovariateAnalysis', 'dplyr', 'plyr')) %dopar% {
  ind = sample(1:length(y), round(length(y)*0.8))
  ind.comp = setdiff(1:length(y), ind)

  obj = cv.glmnet(x[ind,],y[ind], family = 'gaussian')
  coeff = as.matrix(coef.glmnet(obj, s = 'lambda.min'))
  ind.var1 = rownames(coeff)
  ind.var2 = setdiff(rownames(coeff)[coeff != 0], '(Intercept)')

  # Prune model based on least variable of importance
  while(length(ind.var1) > length(ind.var2)){
    obj1 = cv.glmnet(x[ind,ind.var2],y[ind], family = 'gaussian')
    coeff = as.matrix(coef.glmnet(obj1, s = 'lambda.min'))
    ind.var1 = ind.var2
    ind.var2 = setdiff(rownames(coeff)[coeff != 0], '(Intercept)')
  }

  mse = sum((predict(obj1, x[ind.comp,ind.var2], s = 'lambda.min', type = 'response') - y[ind.comp])^2)/length(ind.comp)
  coeff = as.matrix(coef.glmnet(obj1, s = 'lambda.min')) %>%
    rownameToFirstColumn('Coefficients') %>%
    plyr::rename(c('1' = i))
  list(mse = mse, coeff = coeff)
}
boxplot(sapply(glm.model, function(x){x$mse}), main = 'glmnet', ylab = 'mse')
coeff1 = lapply(glm.model, function(x){x$coeff}) %>%
  join_all(type = 'full')
coeff1[is.na(coeff1)] = 0
coeff1 = coeff1[which(rowSums(coeff1[,-(1)] == 0) < 80),]

# Random forest model
rf.model = foreach(i = 1:100,
                .export = c('x','y'),
                .packages = c('ranger', 'CovariateAnalysis', 'dplyr', 'plyr')) %dopar% {
  ind = sample(1:length(y), round(length(y)*0.8))
  ind.comp = setdiff(1:length(y), ind)

  obj = ranger(formula = y ~., data = cbind(data.frame(y = y[ind]), x[ind,]), 
               importance = 'permutation',
               write.forest = TRUE)
  coeff = obj$variable.importance
  ind.var1 = names(coeff)
  ind.var2 = names(coeff)[coeff >= abs(min(coeff))]

  # Prune model based on least variable of importance
  while(length(ind.var1) > length(ind.var2)){
    obj1 = ranger(formula = y ~., data = cbind(data.frame(y = y[ind]), x[ind,ind.var2]),
                  importance = 'permutation',
                  write.forest = TRUE)
    coeff = obj1$variable.importance
    ind.var1 = ind.var2
    ind.var2 = names(coeff)[coeff >= abs(min(coeff))]
  }

  mse = obj1$prediction.error
  coeff = obj1$variable.importance %>%
    rownameToFirstColumn('Coefficients') %>%
    plyr::rename(c('DF' = i))
  list(mse = mse, coeff = coeff)
}
boxplot(sapply(rf.model, function(x){x$mse}), main = 'random forest', ylab = 'mse')
coeff2 = lapply(rf.model, function(x){x$coeff}) %>%
  join_all(type = 'full')
coeff2[is.na(coeff2)] = 0
coeff2 = coeff2[which(rowSums(coeff2[,-(1)] == 0) < 80),]
```

### Store results in synapse
```{r syn.store}
# Create folder to store results in synapse
CODE = Folder(name = 'Predict PMI', parentId ="syn5570291")
CODE = synStore(CODE)

# Store features in synapse (union of both the models)
tmp = setdiff(unique(c(coeff1$Coefficients, coeff2$coefficients)), '(Intercept)')
write.table(tmp, file = 'features.tsv', sep = '\n', row.names = F, quote=F)
obj = File('features.tsv', name = 'Predictors of PMI', parentId = CODE$properties$id)
obj = synStore(obj, used = as.character(ALL_USED_IDs), executed = thisFile, 
               activityName = activityName, activityDescription = activityDescription)
```