### 2016-09-02 code/Rmd/InvestigateCoexpIssue.Rmd
Temp commit

---
title: "Investigate co-expression issue in reprocessed RNASeq"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
# Investigate co-expression issue
# Knitting code (commented)
knit2synapse::knitToFolderEntity(file = "./InvestigateCoexpIssue.Rmd",
                                 parentId = 'syn5569102',
                                 entityName = 'RNASeq Reprocessing Coexpression Quality Control')
```

```{r libload}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(Biobase)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

library(synapseClient)
library(knitr)
library(githubr)

library(Vennerable)
synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```
### Data download
#### Obtain adjusted FPKM/logcpm residual matrix from synapse
```{r data.download, cache = F}
original.residuals.id = c('ROSMAP' = 'syn5581268', 'MAYO' = 'syn6088523', 'MSSM' = 'syn6120103')
all.used.ids = original.residuals.id
original.residuals = lapply(original.residuals.id, downloadFile)
original.residuals.cor = lapply(original.residuals, function(x){cor(t(x[,-(1:2)]))})

h = mapply(function(x,y){ hist(x, main = y) }, original.residuals.cor,  names(original.residuals.cor), SIMPLIFY = F)
writeLines('Adjusted residuals of original data deposited from the individual sites')
plot(h[[1]], main = names(original.residuals.cor)[1])
plot(h[[2]], main = names(original.residuals.cor)[2])
plot(h[[3]], main = names(original.residuals.cor)[3])
```
#### Obtain adjusted residual matrix from reprocessed counts in synapse
```{r data.download2, cache=F}
reproc.residuals.id = c('ROSMAP' = 'syn6114451', 'MAYO' = 'syn6128271', 'MSSM' = 'syn6131415')
all.used.ids = c(all.used.ids, reproc.residuals.id)
reproc.residuals = lapply(reproc.residuals.id, downloadFile)
reproc.residuals.cor = lapply(reproc.residuals, function(x){cor(t(x[,-(1:3)]))})

h = lapply(reproc.residuals.cor, hist)
writeLines('Adjusted residuals of reprocessed data deposited by KKDang')
plot(h[[1]], main = names(reproc.residuals.cor)[1])
plot(h[[2]], main = names(reproc.residuals.cor)[2])
plot(h[[3]], main = names(reproc.residuals.cor)[3])
```
### Extract PC's and find the loading genes in the reprocessed data from each site
#### From the reprocessed data
```{r sample}
reproc.pc.samples = lapply(reproc.residuals, function(x){
  x = x[!duplicated(x$Gene.ID),]
  rownames(x) = x$Gene.ID
  x = x[,-(1:3)]

  PC = prcomp(x, scale. = T, center = T)
  writeLines('Percent variance explained by top 10 PCs are')
  print(((PC$sdev^2/sum(PC$sdev^2)) * 100)[1:10])

  return(PC)
})
pairs(reproc.pc.samples$ROSMAP$rotation[,1:4], main = 'ROSMAP')
pairs(reproc.pc.samples$MSSM$rotation[,1:4], main = 'MSSM')
pairs(reproc.pc.samples$MAYO$rotation[,1:4], main = 'MAYO')
```
##### Find all samples deviating from 2 SD in the first PC
```{r gene}
dev.samples = lapply(reproc.pc.samples, function(x){
  mn = mean(x$rotation[,1]); sdev = sd(x$rotation[,1]);
  zscr = (x$rotation[,1] - mn)/sdev
  print(names(zscr)[abs(zscr) >= 2])
})

reproc.pc.genes = lapply(reproc.residuals, function(x){
  x = x[!duplicated(x$Gene.ID),]
  rownames(x) = x$Gene.ID
  x = x[,-(1:3)]

  PC = prcomp(t(x), scale. = T, center = T)

  writeLines('Percent variance explained by top 10 PCs are')
  print(((PC$sdev^2/sum(PC$sdev^2)) * 100)[1:10])

  return(PC)
})
pairs(reproc.pc.genes$ROSMAP$rotation[,1:4], main = 'ROSMAP')
pairs(reproc.pc.genes$MSSM$rotation[,1:4], main = 'MSSM')
pairs(reproc.pc.genes$MAYO$rotation[,1:4], main = 'MAYO')

##### Find all samples deviating from 2 SD in the first PC
dev.genes = mapply(function(x,y){
  mn = mean(x$rotation[,y]); sdev = sd(x$rotation[,y]);
  zscr = (x$rotation[,y] - mn)/sdev
  genes = (names(zscr)[abs(zscr) >= 2])
}, reproc.pc.genes, c(1,2,1), SIMPLIFY = F)
tmp = Venn(dev.genes)
writeLines('Number of top deviating genes (2 SD)')
plot(tmp, doWeights = F)
writeLines('14 genes that are common to all three data sets are')
genes = intersect(intersect(dev.genes$ROSMAP,dev.genes$MAYO), dev.genes$MSSM)
paste(genes, collapse = ',') %>% print()
```
## Store results in Synapse
```{r syn.store}
parentId = 'syn5569102';
activityName = 'Quality control';

thisFileName <- 'InvestigateCoexpIssue.Rmd'

### Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))

### Store results in synapse
CODE = Folder(name='RNASeq Reprocessing Coexpression Quality Control', parentId = parentId)
CODE = synStore(CODE)

save(list = ls(), file = 'investigate.RData')
# system('zip investigate investigate.RData')
# obj = File('./InvestigateCoexpIssue.Rmd', name = 'RNASeq Reprocessing Coexpression Quality Control', parentId = CODE$properties$id)
# obj = synStore(obj, activityName = activityName, used = as.character(all.used.ids), executed = thisFile)
```