### 2016-09-16 code/Rmd/curateDREAM5DataSets.Rmd
Added markdown to process DREAM5 datasets

---
title: "Format DREAM5 network prediction data for reuse with metanetwork pipeline"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)
library(knitr)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./curateDREAM5DataSets.Rmd",
                                 parentId ="syn5569099",
                                 entityName = 'DREAM5 Networks')
```
#### Initialise libraries
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(synapseClient)
library(knitr)
library(githubr)

library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(githubr) ## Needs the dev branch

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5569099';
activityName = 'Curate DREAM5 networks';
activityDescription = 'Curate DREAM5 data for reuse';

thisFileName <- 'curateDREAM5DataSets.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='synNet')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))

# Create folder to store results
FOLD = Folder(name = 'DREAM5 Networks', parentId = parentId)
FOLD = synStore(FOLD)
```
#### Extract data and network
```{r extractNet, cache=FALSE}
all.folders = synQuery('select * from Folder where parentId == "syn2787212"')
all.expr.obj = dlply(all.folders, .(Folder.id), .fun = function(x){
  all.files = synQuery(paste0('select * from File where parentId == "',x$Folder.id,'"'))

  # Get expression data
  ind.expr = grep('expression', all.files$File.name)
  expr = read.table(synGet(all.files$File.id[ind.expr])@filePath, sep = '\t', header = T, fill=NA)

  # Get gene names
  ind.genes = grep('gene', all.files$File.name)
  genes = read.table(synGet(all.files$File.id[ind.genes])@filePath, sep = '\t', header = F, fill=NA)

  colnames(expr) = genes$V2
  rownames(expr) = paste0('EXPR',1:dim(expr)[1])
  expr = t(expr)

  # Store expression matrix in synapse
  fold.obj = Folder(name = x$Folder.name, parentId = FOLD@properties$id)
  fold.obj = synStore(fold.obj)

  write.table(expr %>% rownameToFirstColumn('GeneNames'), 'Expression.tsv',sep = '\t',row.names = F, quote=F)
  expr.obj = File('Expression.tsv', name = 'Expression Matrix', parentId = fold.obj@properties$id)
  expr.obj = synStore(expr.obj, activityName = activityName, activityDescription = activityDescription,
                      executed = thisFile, used = all.files$File.id)
})
```