### 2016-10-26 code/Rmd/BM36_Compare_All_Data.Rmd
Added markdown to compare all BM36 datasets

---
title: "Compare different versions of BM 36 from MSSM"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./BM36_Compare_All_Data.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 36 Comparison between all versions')
```

### Download all versions of LogCPM from synapse
```{r data.download}
library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(synapseClient)
library(knitr)
library(githubr)

synapseLogin()

lcpm.fromPortal = read.table(synGet('syn7255229')@filePath, row.names = 1, header = T)
lcpm.fromFASTQ = read.table(synGet('syn7255101')@filePath, row.names = 1, header = T)
lcpm.fromBAM = read.table(synGet('syn7255067')@filePath, row.names = 1, header = T)
lcpm.fromSynapseBAM = read.table(synGet('syn7437407')@filePath, row.names = 1, header = T)
```

```{r curate}
Gene.ID = data.frame(Gene.ID = c(rownames(lcpm.fromPortal), 
                                 rownames(lcpm.fromFASTQ), 
                                 rownames(lcpm.fromBAM),
                                 rownames(lcpm.fromSynapseBAM))) %>%
  dplyr::mutate(GeneID = Gene.ID) %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id','position'), sep = '\\.')

tmp = filter(Gene.ID, GeneID %in% rownames(lcpm.fromBAM)) %>% unique
lcpm.fromBAM = WGCNA::collapseRows(lcpm.fromBAM, tmp$ensembl_gene_id, tmp$GeneID)$datETcollapsed

tmp = filter(Gene.ID, GeneID %in% rownames(lcpm.fromFASTQ)) %>% unique
lcpm.fromFASTQ = WGCNA::collapseRows(lcpm.fromFASTQ, tmp$ensembl_gene_id, tmp$GeneID)$datETcollapsed

tmp = filter(Gene.ID, GeneID %in% rownames(lcpm.fromSynapseBAM)) %>% unique
lcpm.fromSynapseBAM = WGCNA::collapseRows(lcpm.fromSynapseBAM, tmp$ensembl_gene_id, tmp$GeneID)$datETcollapsed

genes = intersect(rownames(lcpm.fromSynapseBAM), rownames(lcpm.fromFASTQ)) %>% 
  intersect(rownames(lcpm.fromBAM)) %>%
  intersect(rownames(lcpm.fromPortal))

colnames(lcpm.fromPortal) = gsub('hB_RNA_','BM_36_',colnames(lcpm.fromPortal))
colnames(lcpm.fromSynapseBAM) = gsub('hB_RNA_','BM_36_',colnames(lcpm.fromSynapseBAM))
samples = intersect(colnames(lcpm.fromPortal), colnames(lcpm.fromFASTQ)) %>%
  intersect(colnames(lcpm.fromBAM)) %>%
  intersect(colnames(lcpm.fromSynapseBAM))

lcpm.fromSynapseBAM = lcpm.fromSynapseBAM[genes, samples]
lcpm.fromBAM = lcpm.fromBAM[genes, samples]
lcpm.fromFASTQ = lcpm.fromFASTQ[genes, samples]
lcpm.fromPortal = lcpm.fromPortal[genes, samples]
```

```{r correlate}
# par(mfrow = c(3,1))
cr.por.bam = stats::cor(lcpm.fromPortal, lcpm.fromBAM)
# hist(diag(cr.dep.bam), main = 'Correlation between sinaiDeposit and fromBAM')
# 
cr.por.fastq = stats::cor(lcpm.fromPortal, lcpm.fromFASTQ)
# hist(diag(cr.dep.fastq), main = 'Correlation between sinaiDeposit and fromFASTQ')
# 
cr.por.synbam = stats::cor(lcpm.fromPortal, lcpm.fromSynapseBAM)
# hist(diag(cr.dep.synbam), main = 'Correlation between sinaiDeposit and fromSynapseBAM')
# 
cr.fastq.synbam = stats::cor(lcpm.fromFASTQ, lcpm.fromSynapseBAM)
# hist(diag(cr.bam.fastq), main = 'Correlation between fromBAM and fromFASTQ')

cr.fastq.bam = stats::cor(lcpm.fromFASTQ, lcpm.fromBAM)

tmp = rbind(data.frame(Comp = 'Portal_vs_BAM', values = diag(cr.por.bam)),
            data.frame(Comp = 'Portal_vs_FASTQ', values = diag(cr.por.fastq)),
            data.frame(Comp = 'Portal_vs_synBAM', values = diag(cr.por.synbam)),
            data.frame(Comp = 'FASTQ_vs_synBAM', values = diag(cr.fastq.synbam)),
            data.frame(Comp = 'FASTQ_vs_BAM', values = diag(cr.fastq.bam)))
p = ggplot(tmp, aes(x = Comp, y = values)) + geom_boxplot()
p = p + ylab('Correlation') + xlab('')
p
```