### 2016-11-24 code/Rmd/ROSMAP_Comparison_All_Data.Rmd
Added markdown to compare all versions of ROSMAP

---
title: "Compare different versions of ROSMAP data"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)
library(githubr)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./ROSMAP_Comparison_All_Data.Rmd",
                                 parentId ="syn5570291",
                                 entityName = 'ROSMAP RNASeq Comparisons')
```

### Download all versions of LogCPM from synapse
```{r data.download}
library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(Vennerable)
library(edgeR)
library(geneLenDataBase)

library(synapseClient)
library(knitr)
library(githubr)

synapseLogin()

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r get.data}
fpkm = downloadFile('syn3505720')
rownames(fpkm) = fpkm$tracking_id
fpkm = fpkm[,-c(1:2)]

cnames = data.frame(cnames = colnames(fpkm)) %>%
  dplyr::mutate(cn = cnames) %>%
  separate(cn, c('c1','c2','c3'), sep = '\\_') %>%
  unite(cn, c1, c2)
fpkm = fpkm[,which(!duplicated(cnames$cn))]
colnames(fpkm) = cnames$cn[which(!duplicated(cnames$cn))]

count = list(Lilly = read.table(synGet('syn6112813')@filePath, sep = '\t', quote = '', check.names = F),
             Old = read.table(synGet('syn7211360')@filePath, sep = '\t', quote = '', check.names = F),
             New = read.table(synGet('syn7750270')@filePath, sep = '\t', quote = '', check.names = F))

lcpm = list(Lilly = downloadFile('syn6114447'),
            Old = downloadFile('syn7211581'),
            New = downloadFile('syn7765662'))
```

```{r compute.rpkm}
# Define biomart object
mart <- useMart(host = "jul2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl", biomart = "ENSEMBL_MART_ENSEMBL")

# Get genes to query in biomart
Gene.ID = lapply(count, rownames) %>%
  Reduce(c, .) %>%
  unique %>% 
  data.frame(Gene.ID = .) %>%
  dplyr::mutate(GeneID = Gene.ID) %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id','position'), sep = '\\.') %>% 
  unique

# Query biomart for transcript length
Ensemble2GeneLength <- getBM(attributes = c("ensembl_gene_id", "transcript_length"),
                             filters = "ensembl_gene_id", values = Gene.ID$ensembl_gene_id,
                             mart = mart) %>%
  group_by(ensembl_gene_id) %>%
  summarise(gene.len = median(transcript_length, na.rm = T)) %>%
  left_join(Gene.ID)

# Compute fpkm from counts
all.fpkm = lapply(count, function(x, ens.len){
  len = ens.len$gene.len
  names(len) = ens.len$GeneID
  len = len[rownames(x)]

  x = edgeR::rpkm(x, gene.length = len)
  rs = rowSums(x, na.rm = T) 
  x = x[!is.na(rs) & (rs != 0),]
  return(x)
}, Ensemble2GeneLength)

# Convert lcpm and fpkm matrices to ensemble ids
all.fpkm = lapply(all.fpkm, function(x){
  GeneID = data.frame(Gene.ID = rownames(x)) %>%
    dplyr::mutate(GeneID = Gene.ID) %>%
    tidyr::separate(Gene.ID, c('ensembl_gene_id','position'), sep = '\\.')

  y = WGCNA::collapseRows(x, GeneID$ensembl_gene_id, rownames(x))$datETcollapsed
  return(y)
})

lcpm = lapply(lcpm, function(x){
  x = x[!duplicated(x$Gene.ID),]
  rownames(x) = x$Gene.ID
  y = WGCNA::collapseRows(x[,-c(1:3)], x$ensembl_gene_id, x$Gene.ID)$datETcollapsed
  return(y)
})

GeneID = data.frame(Gene.ID = rownames(fpkm)) %>%
  dplyr::mutate(GeneID = Gene.ID) %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id','position'), sep = '\\.')
all.fpkm$Portal = WGCNA::collapseRows(fpkm, GeneID$ensembl_gene_id, rownames(fpkm))$datETcollapsed

###################################################
# Find unique IDs and samples across all data
GeneID = lapply(all.fpkm, rownames) %>%
  Reduce(intersect, .) %>%
  intersect(lapply(lcpm, rownames) %>%
              Reduce(intersect,.)) %>%
  unique

SampleID = lapply(all.fpkm, colnames) %>%
  Reduce(intersect, .) %>%
  intersect(lapply(lcpm, colnames) %>%
              Reduce(intersect, .)) %>%
  unique

all.fpkm = lapply(all.fpkm, function(x, GeneID, SampleID){
  x = x[GeneID, SampleID]
}, GeneID, SampleID)

lcpm = lapply(lcpm, function(x, GeneID, SampleID){
  x = x[GeneID, SampleID]
}, GeneID, SampleID)
```

```{r correlate.fpkm, fig.height=5, fig.width=8}
cr1 = stats::cor(all.fpkm[['Lilly']],all.fpkm[['Old']], use = 'pairwise.complete.obs') 
all.cr = data.frame(data = 'Lilly_vs_Old', values = diag(cr1))

cr2 = stats::cor(all.fpkm[['Lilly']],all.fpkm[['New']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Lilly_vs_New', values = diag(cr2)))

cr3 = stats::cor(all.fpkm[['Old']],all.fpkm[['New']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Old_vs_New', values = diag(cr3)))

cr4 = stats::cor(all.fpkm[['Lilly']],all.fpkm[['Portal']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Lilly_vs_Portal', values = diag(cr4)))

cr5 = stats::cor(all.fpkm[['Old']],all.fpkm[['Portal']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Old_vs_Portal', values = diag(cr5)))

cr6 = stats::cor(all.fpkm[['New']],all.fpkm[['Portal']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'New_vs_Portal', values = diag(cr6)))

writeLines('')
p = ggplot(all.cr, aes(x = data, y = values)) + geom_boxplot()
p = p + ylab('Correlation') + xlab('') + ggtitle('Correlation between FPKM values')
p
```

```{r correlate.lcpm, fig.height=5, fig.width=5}
cr1 = stats::cor(lcpm[['Lilly']],lcpm[['Old']], use = 'pairwise.complete.obs') 
all.cr = data.frame(data = 'Lilly_vs_Old', values = diag(cr1))

cr2 = stats::cor(lcpm[['Lilly']],lcpm[['New']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Lilly_vs_New', values = diag(cr2)))

cr3 = stats::cor(lcpm[['Old']],lcpm[['New']], use = 'pairwise.complete.obs') 
all.cr = rbind(all.cr, data.frame(data = 'Old_vs_New', values = diag(cr3)))

writeLines('')
p = ggplot(all.cr, aes(x = data, y = values)) + geom_boxplot()
p = p + ylab('Correlation') + xlab('') + ggtitle('Correlation between lcpm values')
p
```

```{r sample.dev, results='asis'}
PC = lapply(lcpm, function(x){ prcomp(x, center = T, scale. = T)})

writeLines('Percent variance explained by first 10 PCs')
tmp = sapply(PC, function(x){
  pv = (x$sdev^2/sum(x$sdev^2))*100
  pv[1:10]
}) 
rownames(tmp) = paste0('PC',1:10)
kable(tmp)

writeLines('Deviating samples (2SD from PC1 median) in all three cases')
dev.samples = lapply(PC, function(x){ 
  lb = median(x$rotation[,1]) - 2*sd(x$rotation[,1]) 
  ub = median(x$rotation[,1]) + 2*sd(x$rotation[,1]) 
  rownames(x$rotation)[c(which(x$rotation[,1]<lb), which(x$rotation[,1] > ub))]
})
plot(Venn(dev.samples), doWeights = F)
tmp = mapply(function(x, y){writeLines(y); writeLines(paste(x, collapse = ','))}, dev.samples, names(dev.samples))
```