### 2017-02-02 code/Rmd/ROSMAP_RNASeq_Reprocessed_TMM_VOOM_CellProp.Rmd
Added markdown to deconvolute cell typr fraction in reprocessed RNASeq from ROSMAP

---
title: "Markdown to estimate the abundance of brain cell types with the reprocessed counts from ROSMAP cohort (with TMM-VOOM normalisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./ROSMAP_RNASeq_Reprocessed_TMM_VOOM_CellProp.Rmd",
                                 parentId = "syn5570291",
                                 entityName = 'ROSMAP Cell Type Abundance with Reprocessed RNASeq')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(Biobase)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(xlsx)

library(synapseClient)
library(knitr)
library(githubr)

library(parallel)
library(doParallel)
library(doMC)
library(foreach)

cl = makeCluster(detectCores() - 2)
registerDoParallel(cl)

synapseLogin()

# Source external files
source('../R/CIBERSORT.R')

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570291';
activityName = 'Cell type abundance in ROSMAP';
activityDescription = 'Cell type abundance in ROSMAP with TMM-VOOM normalised reprocessed RNASeq counts';

thisFileName <- 'ROSMAP_RNASeq_Reprocessed_TMM_VOOM_CellProp.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='celltype')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn7750270';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID)
COUNT <- read.table(COUNT_OBJ@filePath,header=T,sep='\t',check.names = F)

# Get clinical metadata
METADATA.CLINICAL_ID <- 'syn3191087'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID
METADATA.CLINICAL_OBJ <- synGet(METADATA.CLINICAL_ID)
METADATA.CLINICAL <- read.table(METADATA.CLINICAL_OBJ@filePath,sep=',',header=T)

# Get technical covariates
METADATA.TECH_ID <- 'syn4300313'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.TECH_ID
METADATA.TECH_OBJ <- synGet(METADATA.TECH_ID)
METADATA.TECH <- read.table(METADATA.TECH_OBJ@filePath,sep='\t',header=T)

# Fix error in technical covariates data
KEY_ID <- 'syn3382527'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = KEY_ID  
KEY <- read.csv(synGet(KEY_ID)@filePath) %>% 
  filter(mrna_data == 1) %>%
  dplyr::select(projid, mrna_id) %>%
  separate(mrna_id, c('a','b','batch'), sep = '_') %>% 
  unite(Sampleid, a, b) %>%
  dplyr::select(-batch) %>%
  unique

# Match technical and clinical covariates
METADATA <- METADATA.TECH %>% 
  dplyr::select(-projid) %>%
  dplyr::left_join(KEY) %>%
  dplyr::left_join(METADATA.CLINICAL)

# Pick higher quality RIN batch for sample 492_120515
METADATA <- METADATA %>%
  group_by(Sampleid) %>%
  top_n(1, RINcontinuous)
```

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  ungroup %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi)) %>%
  filter(!is.na(Sampleid)) %>%
  filter(!(Sampleid %in% c('380_120503'))) %>%
  as.data.frame()
```

```{r preprocess.data1,cache=TRUE}
# Match covariates to expression data
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))
removedIDs = setdiff(colnames(COUNT), METADATA$Sampleid)

COUNT = COUNT[,indToRetain]

rownames(METADATA) = METADATA$Sampleid
METADATA = METADATA[indToRetain,]
```
Dorsolateral prefrontal cortex of `r dim(COUNT)[2]` subjects from the ROS and MAP cohorts are used for the analysis. Following sample are removed `r paste(removedIDs, collapse = ',')`

### Get Covariates
```{r covariates.clustering, cache=TRUE}
#"braaksc","ceradsc","cts_mmse30_first_ad_dx","cts_mmse30_lv"
FactorCovariates <- c("Batch", "msex", "race", "spanish", "cogdx")
ContCovariates <- c("RINcontinuous","age_first_ad_dx","age_death","age_at_visit_max","pmi","educ")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)

# Add in RIN^2 values
COVARIATES$RINcontinuous2 = COVARIATES$RINcontinuous^2
ContCovariates = c(ContCovariates, 'RINcontinuous2')
```

### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

Remove genes that have less than 1 cpm counts in at least 50% of samples per cogdx
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(cogdx), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% unique()

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Normalisation (with NULL)
Initial normalisation usign voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=T)

# Set gene counts in specific samples that are deviating 3 sd from other samples to NA
log.mat = apply(VOOM.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  return((x < (mn-3*std.dev)) | (x > (mn+3*std.dev)))
}) %>% t
PROCESSED_COUNTS$filteredExprMatrix[log.mat] = NA

# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=T)
```

### Get cell type specific signatures and expression data from synapse
```{r get.signatures}
ALL_USED_IDs = c(ALL_USED_IDs, 'syn8068511', 'syn8068475')
# Signatures from Zhang,Y et al., 2014
sig.zhang = downloadFile('syn8068511') %>%
  dlply(.(Set), .fun = function(x){return(unique(x$Gene))})

# Signatures from Darmanis, S et al., 2015
sig.dar.obj = synGet('syn8068475')  
sig.dar.unbiased = read.xlsx(sig.dar.obj@filePath, sheetName = 'Unbiased') %>%
  lapply(function(x){as.character(unique(x))})
sig.dar.unbiased = sig.dar.unbiased[1:10]

# Extract all signatures
sig = list()
sig$OPC = unique(c(sig.zhang$OPC, sig.dar.unbiased$Cluster.1))
sig$Oligodendrocyte = unique(c(sig.zhang$MyelinOligos, sig.zhang$NewOligos, sig.dar.unbiased$Cluster.3))
sig$Neuron = unique(c(sig.zhang$Neuron, sig.dar.unbiased$Cluster.6, sig.dar.unbiased$Cluster.7,
                      sig.dar.unbiased$Cluster.10, sig.dar.unbiased$Cluster.2))
sig$Endothelial = unique(c(sig.zhang$Endothelial, sig.dar.unbiased$Cluster.8))
sig$Astrocyte = unique(c(sig.zhang$Astrocyte, sig.dar.unbiased$Cluster.4))
sig$Microglia = unique(c(sig.zhang$Astrocyte, sig.dar.unbiased$Cluster.5))

sig = ldply(sig, .fun = function(x){as.data.frame(x)})
colnames(sig) = c('CellType','Gene.Name')

# Get reference expression data from Darmanis., S et al., 2015
ALL_USED_IDs = c(ALL_USED_IDs, 'syn8077142', 'syn8077191')
ref.expr = downloadFile('syn8077142')
rownames(ref.expr) = ref.expr$hgnc_symbol
ref.expr$hgnc_symbol = NULL

ref.cellTypes = downloadFile('syn8077191') %>%
  dplyr::select(SampleID, CellType)

ref.expr1 = list()
for(celltype in unique(ref.cellTypes$CellType)){
  ids = ref.cellTypes$SampleID[ref.cellTypes$CellType == celltype]
  ref.expr1[[celltype]] = log2(rowSums(2^ref.expr[,ids], na.rm = T)) %>%
    rownameToFirstColumn('hgnc_symbol') %>%
    plyr::rename(c('DF' = celltype))
}
ref.expr1 = join_all(ref.expr1)
```

```{r get.gene.symbols, cache=TRUE, eval=T}
backgroundGenes = data.frame(gene.id = rownames(VOOM.GENE_EXPRESSION$E)) %>%
  dplyr::mutate(ensembl_gene_id = gene.id) %>%
  tidyr::separate(ensembl_gene_id, c('ensembl_gene_id', 'position'), sep = '\\.')

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", 
                       values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

# Add gene names
EXPR <- VOOM.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('gene.id') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  filter(!duplicated(gene.id))
rownames(EXPR) = EXPR$gene.id

EXPR1 = WGCNA::collapseRows(EXPR %>% dplyr::select(-ensembl_gene_id, -position, -hgnc_symbol, -gene.id), 
                            EXPR$hgnc_symbol, 
                            EXPR$gene.id)$datETcollapsed

ind = intersect(rownames(EXPR1), ref.expr1$hgnc_symbol)
EXPR1 = EXPR1 %>% 
  rownameToFirstColumn('hgnc_symbol') %>%
  filter(hgnc_symbol %in% ind)
ref.expr1 = ref.expr1 %>%
  filter(hgnc_symbol %in% ind)

write.table(ref.expr1, file = 'ref.expr.txt', sep = '\t', quote=F, row.names = F)
write.table(EXPR1[,1:20], file = 'mixture.expr.txt', sep = '\t', quote=F, row.names = F)

cellprop = CIBERSORT('ref.expr.txt', 'mixture.expr.txt', perm = 100) # Obtain CIBERSORT.R function from http://cibersort.stanford.edu/
stopCluster(cl)
```

```{r plot.cellprop}
cellprop1 = as.data.frame(cellprop)
cellprop1$cogdx = COVARIATES[rownames(cellprop), 'cogdx']
cellprop1 = cellprop1 %>% 
  rownameToFirstColumn('SampleID') %>%
  dplyr::select(-`P-value`,-Correlation, -RMSE) %>%
  tidyr::gather(Celltype, Fraction, -SampleID, -cogdx) %>%
  dplyr::arrange(desc(cogdx))

p = ggplot(cellprop1, aes(x = SampleID, y = Fraction, fill = Celltype, color = cogdx)) + geom_bar(stat = 'identity')
p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
# p = p + facet_grid(.~cogdx)
p

p = ggplot(cellprop1, aes(x = Celltype, y = Fraction, color = cogdx)) + geom_boxplot()
p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p
```
### Synapse store
```{r syn.store, cache=TRUE}
CODE = Folder(name = 'ROSMAP Cell Type Abundance with Reprocessed RNASeq', parentId = parentId)
CODE = synStore(CODE)

cellprop %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_CellFraction_Reproc.tsv', row.names = F, sep = '\t', quote = F)
obj = File('ROSMAP_CellFraction_Reproc.tsv', name = 'ROSMAP Cell Fraction', parentId = CODE$properties$id)
obj = synStore(obj, activityName = activityName, activityDescription = activityDescription,
               used = ALL_USED_IDs, executed = thisFile)
```