### 2017-02-21 code/Rmd/combineResults.Rmd
Added markdown to combine resulat and perform enrichment analysis of pathways

---
title: "Combine differential expression and celltype fraction results from all brain banks"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./combineResults.Rmd", 
                                 parentId ="syn5569102",
                                 entityName = 'All Differential Expression Results')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")

# Load required libraries
library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(RColorBrewer)
library(ggplot2)
library(ComplexHeatmap)

library(RankProd)
library(githubr)

library(Vennerable)
library(colorfulVennPlot)

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Get covariates/sample metadata from synapse
```{r covariates}
covariates.ids = c(ROSMAP = 'syn8018344',
                   MSBB = 'syn8073656',
                   MAYO = 'syn8028561')
covariates = lapply(covariates.ids, function(id){
  x = downloadFile(id)
  colnames(x)[1] = 'SampleID'
  return(x)
})

covariates$ROSMAP = dplyr::filter(covariates$ROSMAP, cogdx %in% c(1,4)) %>%
  dplyr::mutate(Dx = factor(cogdx, levels = c(1,4), labels = c('1' = 'NCI', '4' = 'AD')),
                Study = 'ROSMAP',
                BrainRegion = 'DLPFC') %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MSBB = dplyr::filter(covariates$MSBB, BrodmannArea.Dx %in% c('BM10.SD', 'BM22.SD', 'BM36.SD', 'BM44.SD', 'BM36.ND', 'BM10.ND', 'BM44.ND', 'BM22.ND')) %>%
  dplyr::select(SampleID, BrodmannArea.Dx) %>%
  tidyr::separate(BrodmannArea.Dx, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MSSM',
                Dx = factor(Dx, levels = c('ND','SD')),
                BrainRegion = factor(BrainRegion, labels = c('FP', 'STG', 'PHG', 'IFG'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MAYO = dplyr::filter(covariates$MAYO, BrainRegion.Diagnosis %in% c('TCX.AD', 'TCX.Control', 'CER.AD', 'CER.Control')) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MAYO',
                Dx = factor(Dx, levels = c('Control','AD')),
                BrainRegion = factor(BrainRegion, levels = c('CER', 'TCX'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

diff.exp.ids = c(ROSMAP = 'syn8018358',
                 MSBB = 'syn8073678',
                 MAYO = 'syn8028572')
diff.exp = lapply(diff.exp.ids, downloadFile)

diff.exp$ROSMAP = filter(diff.exp$ROSMAP, Comparison == 'cogdx1-cogdx4')
diff.exp$MSBB = filter(diff.exp$MSBB, Comparison == 'ND-SD')
diff.exp$MAYO = filter(diff.exp$MAYO, Comparison == 'Control-AD')

celltype.fraction.ids = c(ROSMAP = 'syn8124520',
                      MSBB = 'syn8146376',
                      MAYO = 'syn8135250')
celltype.fraction = lapply(celltype.fraction.ids, downloadFile) 
```

```{r anlz.celltype}
cell.frac = mapply(function(cov, frc){
  x = filter(frc, SampleID %in% cov$SampleID) %>%
    dplyr::select(-`P-value`,-RMSE, -Correlation) %>%
    tidyr::gather(celltype, fraction, -SampleID) %>%
    left_join(cov)
}, covariates, celltype.fraction, SIMPLIFY = F) %>%
  rbindlist(use.names = T, fill = T) %>%
  dplyr::mutate(Dx = factor(Dx, levels = c('Control', 'NCI', 'ND', 'SD', 'AD')))

head(cell.frac)

p = ggplot(cell.frac, aes(x = BrainRegion, y = fraction, color = Dx)) + geom_boxplot()
p = p + facet_grid(celltype~., scales = 'free_y')
p = p + xlab('Brain Region') + ylab('Fraction')
p = p + theme(axis.text.x = element_text(hjust = 1, angle = 90))
p
```

```{r anlz.diff.exp}
tmp = lapply(diff.exp, function(x){
  x = filter(x, adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) 
  x$Direction = 'NONE'
  x$Direction[x$logFC > 0] = 'UP'
  x$Direction[x$logFC < 0] = 'DOWN'
  x = x %>% group_by(Study, Region, Direction) %>%
    summarise(count = length(unique(gene.id)))
}) %>%
  rbindlist() %>%
  spread(Direction, count)

tmp = diff.exp %>% 
  rbindlist(use.names = T, fill = T) 
tmp$Direction = 'NONE'
tmp$Direction[x$logFC > 0] = 'UP'
tmp$Direction[x$logFC < 0] = 'DOWN'

p = ggplot(tmp, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p
```