### 2017-02-21 code/Rmd/enrichmentAnalysis.Rmd

---
title: "Enrichment analysis of differential expression results from AMP-AD (Control vs AD, reprocessed RNASeq)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./enrichmentAnalysis.Rmd", 
                                 parentId ="syn5569102",
                                 entityName = 'Enrichment Analysis of Control vs AD from all three brain banks (Reprocessed RNASeq)')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")

# Load required libraries
library(CovariateAnalysis) # install from th1vairam repo using devtools::install_github('th1vairam/CovariateAnalysis@devtools')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(RColorBrewer)
library(ggplot2)
library(ComplexHeatmap)
library(biomaRt)
library(Matrix)
library(org.Hs.eg.db)

library(limma)
library(edgeR)
library(RankProd)
library(githubr)

library(Vennerable)
library(colorfulVennPlot)

synapseLogin()

library(doParallel)
library(foreach)
library(doParallel)

clust = makeCluster(detectCores() - 2)
registerDoParallel(clust)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r fxns}
# Function to perform Fishers enrichment analysis
fisherEnrichment <- function(genesInSignificantSet, # A character vector of differentially expressed or some significant genes to test
                             genesInGeneSet, # A character vector of genes in gene set like GO annotations, pathways etc...
                             genesInBackground # Background genes that are 
){
  genesInSignificantSet = intersect(genesInSignificantSet, genesInBackground) # back ground filtering
  genesInNonSignificantSet = base::setdiff(genesInBackground, genesInSignificantSet)
  genesInGeneSet = intersect(genesInGeneSet, genesInBackground) # back ground filtering
  genesOutGeneSet = base::setdiff(genesInBackground,genesInGeneSet)

  pval = fisher.test(
    matrix(c(length(intersect(genesInGeneSet, genesInSignificantSet)),             
             length(intersect(genesInGeneSet, genesInNonSignificantSet)),
             length(intersect(genesOutGeneSet, genesInSignificantSet)),
             length(intersect(genesOutGeneSet, genesInNonSignificantSet))), 
           nrow=2, ncol=2), alternative="greater")
  OR = (length(intersect(genesInGeneSet, genesInSignificantSet)) * length(intersect(genesOutGeneSet, genesInNonSignificantSet))) / (length(intersect(genesInGeneSet, genesInNonSignificantSet)) * length(intersect(genesOutGeneSet, genesInSignificantSet))) %>% as.numeric()
  return(data.frame(pval = pval$p.value,
                    ngenes = length(genesInGeneSet),
                    noverlap = length(intersect(genesInGeneSet, genesInSignificantSet)),
                    Odds.Ratio = OR,
                    Genes = paste(intersect(genesInGeneSet, genesInSignificantSet), collapse = '|')))}
```

### Get covariates/sample metadata from synapse
```{r covariates}
covariates.ids = c(ROSMAP = 'syn8018344',
                   MSBB = 'syn8073656',
                   MAYO = 'syn8028561')
covariates = lapply(covariates.ids, function(id){
  x = downloadFile(id)
  colnames(x)[1] = 'SampleID'
  return(x)
})

covariates$ROSMAP = dplyr::filter(covariates$ROSMAP, cogdx %in% c(1,4)) %>%
  dplyr::mutate(Dx = factor(cogdx, levels = c(1,4), labels = c('1' = 'NCI', '4' = 'AD')),
                Study = 'ROSMAP',
                BrainRegion = 'DLPFC') %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MSBB = dplyr::filter(covariates$MSBB, BrodmannArea.Dx %in% c('BM10.SD', 'BM22.SD', 'BM36.SD', 
                                                                        'BM44.SD', 'BM36.ND', 'BM10.ND', 
                                                                        'BM44.ND', 'BM22.ND')) %>%
  dplyr::select(SampleID, BrodmannArea.Dx) %>%
  tidyr::separate(BrodmannArea.Dx, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MSSM',
                Dx = factor(Dx, levels = c('ND','SD')),
                BrainRegion = factor(BrainRegion, labels = c('FP', 'STG', 'PHG', 'IFG'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MAYO = dplyr::filter(covariates$MAYO, BrainRegion.Diagnosis %in% c('TCX.AD', 'TCX.Control', 'CER.AD', 'CER.Control')) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MAYO',
                Dx = factor(Dx, levels = c('Control','AD')),
                BrainRegion = factor(BrainRegion, levels = c('CER', 'TCX'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)
```

### Get differential expression data from synapse
```{r diffexp}
diff.exp.ids = c(ROSMAP = 'syn8018358',
                 MSBB = 'syn8073678',
                 MAYO = 'syn8028572')
diff.exp = lapply(diff.exp.ids, downloadFile)

diff.exp$ROSMAP = filter(diff.exp$ROSMAP, Comparison == 'cogdx1-cogdx4')
diff.exp$MSBB = filter(diff.exp$MSBB, Comparison == 'ND-SD')
diff.exp$MAYO = filter(diff.exp$MAYO, Comparison == 'Control-AD')

tmp = diff.exp %>% 
  rbindlist(use.names = T, fill = T) 
tmp$Direction = 'NONE'
tmp$Direction[tmp$logFC >= log2(1.2) & tmp$adj.P.Val <= 0.05] = 'UP'
tmp$Direction[tmp$logFC <= -log2(1.2) & tmp$adj.P.Val <= 0.05] = 'DOWN'

writeLines('Number of differentially expressed genes at an FDR of 0.05 and fold change of 1.2 are')
tmp %>%
  group_by(Study, Region, Direction, Comparison) %>%
  summarise(count = n()) %>%
  spread(Direction, count) %>%
  kable

pl = list()
for(region in c('ROSMAP', 'MSBB', 'MAYO')){
  p = ggplot(tmp %>% filter(Study == region), aes(x = logFC, y = -log10(adj.P.Val), color = Direction)) + geom_point()
  p = p + scale_color_manual(values = c('green','grey','red'))
  p = p + facet_grid(Study~.+Region) + theme(legend.position = 'top')
  pl[[region]] = p
}
multiplot(plotlist = pl, cols = 1)
```

### Get hgnc_symbol mapping from biomaRt
```{r biomart}
# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = "dec2016.archive.ensembl.org")

# Get background gene ids
backgroundGenes = sapply(diff.exp, function(x) x$ensembl_gene_id) %>%
  Reduce(c,.)

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = backgroundGenes,
                       mart = mart)

Entrez2Ensembl <- as.data.frame(org.Hs.egENSEMBL2EG)
Entrez2HGNC <- as.data.frame(org.Hs.egSYMBOL2EG)
HGNC2Ensembl <- left_join(Entrez2HGNC, Entrez2Ensembl)
HGNC2Ensembl = HGNC2Ensembl[grep('^LOC',HGNC2Ensembl$symbol),c('ensembl_id', 'symbol')] %>%
  filter(!is.na(ensembl_id), !is.na(symbol)) %>%
  dplyr::rename(ensembl_gene_id = ensembl_id, hgnc_symbol = symbol)

Ensemble2HGNC = rbindlist( list(Ensemble2HGNC, HGNC2Ensembl), use.names = T, fill = T) %>%
  unique
```

### Pathway enrichment analysis
```{r enrich.all, fig.height=16, fig.width=10}
# Find genes that are differentially expressed in each brain region
diff.exp.gl = lapply(diff.exp, function(x){
  dplyr::filter(x, abs(logFC) >= log2(1.2) & adj.P.Val <= 0.05 & sign(CI.L) == sign(CI.R)) %>%
    dplyr::select(Region, hgnc_symbol) %>%
    unique
}) %>%
  rbindlist(use.names = T, fill = T) %>%
  dlply(.(Region), .fun = function(x){unique(x$hgnc_symbol)})

# Get Background genes
backgroundGenes = Ensemble2HGNC$hgnc_symbol[Ensemble2HGNC$ensembl_gene_id %in% backgroundGenes] %>% unique()

#### Get gene sets ####
# Download enrichr gene sets from synapse
GL_OBJ = synGet('syn4867851')
ALL_USED_IDs = c(ALL_USED_IDs, GL_OBJ$properties$id)
load(GL_OBJ@filePath)

gsets = c("BioCarta_2015", "GO_Biological_Process", "Reactome_2015", "TargetScan_microRNA", "WikiPathways_2015")
GeneSets.Enrichr = GeneSets[gsets]

# Download AD related gene sets from synapse
GL_OBJ = synGet('syn4893059');
ALL_USED_IDs = c(ALL_USED_IDs, GL_OBJ$properties$id)
load(GL_OBJ@filePath)

GeneSets.CM = list(CellTypeMarkers = GeneSets[grep('Zhang', names(GeneSets))],
                   AD = GeneSets['AD:GeneticLoci'])
GeneSets = c(GeneSets.Enrichr, GeneSets.CM)

# Filter gene list
GeneSets = filterGeneSets(GeneSets, backgroundGenes, minSize = 3, maxSize = 5000)

# Enrichment Analysis
enrichResults = plyr::ldply(diff.exp.gl, .fun = function(gs, geneSets, backgroundGenes, fisherEnrichment){
  tmp = plyr::ldply(geneSets, .fun = function(x, gs, backgroundGenes, fisherEnrichment){
    plyr::ldply(x, .fun = fisherEnrichment, gs, backgroundGenes, .id = 'Name', .parallel = T)
  }, gs, backgroundGenes, fisherEnrichment, .id = 'Category', .parallel = T, .paropts = list(.packages = c('dplyr'))) %>%
    dplyr::mutate(fdr = p.adjust(pval, method = 'bonferroni'))
}, GeneSets, backgroundGenes, fisherEnrichment, .parallel = T, .paropts = list(.packages = c('dplyr', 'plyr')))

# Filter results
tmp1 = filter(enrichResults, fdr <= 0.05)
```

### Store results in synapse
```{r synapse.store, include=FALSE, cache=FALSE}
thisFileName <- 'enrichmentAnalysis.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='metaAnal')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))

# Store consolidated data in synapse
CODE = Folder(name = 'Pathway Enrichment Analysis of Control vs AD Data (all brain banks)', parentId = "syn5569102")
CODE = synStore(CODE)

# Store enrichment analysis in synapse
write.table(tmp1, file = 'enrichmentResults.tsv', sep = '\t', row.names=F, quote=F)
OBJ = File('enrichmentResults.tsv', 
           name = 'Pathway Enrichment Results (AD and Control)', 
           parentId = CODE$properties$id)
OBJ = synStore(OBJ,
               used = as.character(ALL_USED_IDs),
               executed = thisFile,
               activityName = 'Pathway enrichment analysis')
stopCluster(cl)
```