### 2017-05-07 code/Rmd/combineAllRNASeqData.Rmd
Minor modifications

---
title: "Combine filtered data from all three studies in AMP-AD for metanalysis"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./combineAllRNASeqData.Rmd", 
                                 parentId ="syn9622393",
                                 entityName = "All RNASeq data combined")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")

# Load required libraries
library(CovariateAnalysis) # install from th1vairam repo using devtools::install_github('th1vairam/CovariateAnalysis@devtools')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn9622393';
activityName = 'Combine Data';
activityDescription = 'Combine all RNASeq and covariates data from all three brain banks';

thisFileName <- 'combineAllRNASeqData.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='metaAnal')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Get filtered covariates/sample metadata from synapse
```{r covariates, fig.height=6, fig.width=10}
covariates.ids = c(ROSMAP = 'syn8456631',
                   MSBB = 'syn8484996',
                   MAYO = 'syn8466814')
all.used.ids = as.character(covariates.ids)
covariates = lapply(covariates.ids, downloadFile)

# Modify ROSMAP covariates
rosmap.key = downloadFile('syn3382527') %>%
  dplyr::select(projid, mrna_id) %>%
  tidyr::separate(mrna_id, c('a','b','c'), sep = '_') %>%
  tidyr::unite(SampleID, a,b, sep = '_') %>%
  dplyr::rename(Donor_ID = projid)
all.used.ids = c(all.used.ids, 'syn3382527')

covariates$ROSMAP = covariates$ROSMAP %>%
  left_join(rosmap.key) %>%
  dplyr::mutate(Gender = factor(msex, levels = c(0,1), labels = c('0' = 'FEMALE', '1' = 'MALE')),
                RIN = RINcontinuous,
                AAD = age_death,
                PMI = pmi,
                BrainRegion = 'DLPFC',
                Study = 'ROSMAP') %>%
  dplyr::select(SampleID, Donor_ID, Study, BrainRegion, Diagnosis, Batch, Gender, RIN, AAD, PMI, PCT_PF_READS_ALIGNED,
                PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)

# Modify MSBB covariates
covariates$MSBB = covariates$MSBB %>%
  dplyr::mutate(Gender = factor(SEX, levels = c('F', 'M'), labels = c('F' = 'FEMALE', 'M' = 'MALE')),
                Batch = batch,
                PMI = PMI/60,
                AAD = AOD,
                Donor_ID = individualIdentifier,
                Study = 'MSBB') %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::select(SampleID, Donor_ID, Study, BrainRegion, Diagnosis, Batch, Gender, RIN, AAD, PMI, PCT_PF_READS_ALIGNED,
                PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)

# Modify MAYO covariates
covariates$MAYO = covariates$MAYO %>%
  dplyr::mutate(Gender = factor(Gender, levels = c('F', 'M'), labels = c('F' = 'FEMALE', 'M' = 'MALE')),
                Batch = FLOWCELL,
                AAD = AgeAtDeath,
                Study = factor(Source, 
                               levels = c('BannerSunHealth_TomBeach', 'MayoBrainBank_Dickson'),
                               labels = c('BannerSunHealth_TomBeach' = 'BANNER', 'MAYO'))) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::select(SampleID, Donor_ID, Study, BrainRegion, Diagnosis, Batch, Gender, RIN, AAD, PMI, PCT_PF_READS_ALIGNED,
                PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)

### All plots
covariates1 = rbindlist(covariates, use.names = T, fill = T) %>%
  dplyr::mutate(BrainRegion = factor(BrainRegion, levels = c('DLPFC', 'IFG', 'STG', 'FP', 'PHG', 'CER', 'TCX')))

plotlist = list()
p = ggplot(covariates1, aes(x = Diagnosis, y = PMI, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = AAD, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = RIN, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = PCT_INTERGENIC_BASES, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = PCT_INTRONIC_BASES, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = PCT_RIBOSOMAL_BASES, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = PCT_CODING_BASES, color = BrainRegion)) + geom_boxplot()
p

p = ggplot(covariates1, aes(x = Diagnosis, y = PCT_PF_READS_ALIGNED, color = BrainRegion)) + geom_boxplot()
p
```

### Get raw reprocessed counts from synapse
```{r counts}
# Download sample counts
counts.id = c(ROSMAP = 'syn8691134',
              MSBB = 'syn8691099',
              MAYO1 = 'syn8690799', 
              MAYO2 = 'syn8690904')
all.used.ids = c(all.used.ids, as.character(counts.id))
counts = lapply(counts.id, function(id){
  tmp = read.table(synGet(id)@filePath, header=T, sep='\t', check.names = F, row.names = 1)
})

covariates$MAYO1 = covariates$MAYO
covariates$MAYO2 = covariates$MAYO
counts = mapply(function(cnt, cov){
  cnt[,intersect(colnames(cnt), cov$SampleID)] %>%
    CovariateAnalysis::rownameToFirstColumn('Gene.ID')
}, counts, covariates[names(counts)], SIMPLIFY = F) %>%
  plyr::join_all(type = 'full')
```

### Store results in synapse
```{r synapse.store, include=FALSE, cache=FALSE}
# Store consolidated data in synapse
CODE = Folder(name = 'All RNASeq data combined', parentId = parentId)
CODE = synStore(CODE)

# Store covariates in synapse
write.table(covariates1, file = 'allRNASeqCovariates.tsv', sep = '\t', row.names=F, quote=F)
OBJ = File('allRNASeqCovariates.tsv', 
           name = 'All RNASeq covariates', 
           parentId = CODE$properties$id)
OBJ = synStore(OBJ,
               used = all.used.ids,
               executed = thisFile,
               activityName = activityName,
               activityDescription = activityDescription)

# Store raw counts in synapse
write.table(counts, file = 'allReprocessedCounts.tsv', sep = '\t', row.names=F, quote=F)
OBJ = File('allReprocessedCounts.tsv', 
           name = 'All RNASeq reprocessed raw counts', 
           parentId = CODE$properties$id)
OBJ = synStore(OBJ,
               used = all.used.ids,
               executed = thisFile,
               activityName = activityName,
               activityDescription = activityDescription)
```
### Source code in github
[Source Code](`r thisFile`)