### 2017-05-14 code/Rmd/studyLevelDiffExpAnlz.Rmd
Added markdown to perform differential expression in a study specific manner

---
title: "Study specific differential expression analysis (all 3 studies/one study at a time)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./studyLevelDiffExpAnlz.Rmd", 
                                 parentId ="syn9622393",
                                 entityName = "Differential Expression Analysis (per study)")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")

# Load required libraries
library(CovariateAnalysis) # install from th1vairam repo using devtools::install_github('th1vairam/CovariateAnalysis@devtools')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(limma)
library(edgeR)
library(cqn)
library(knitr)
library(psych)

synapseLogin()

library(foreach)
library(doParallel)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn9622393';
activityName = 'Differential expression analysis';
activityDescription = 'Differential expression analysis in all 7 brain regions performed in a study specific manner';

thisFileName <- 'studyLevelDiffExpAnlz.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='metaAnal')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Get merged covariates and raw counts from synapse
```{r download.data, fig.height=6, fig.width=10}
# Get covariates
covariates = downloadFile('syn9765937') %>%
  dplyr::filter(Diagnosis %in% c('AD', 'CONTROL')) %>%
  dplyr::mutate(Diagnosis.Gender = paste(Diagnosis, Gender, sep = '.'))
all.used.ids = 'syn9765625'

kable(table(covariates$BrainRegion, covariates$Diagnosis))
kable(table(covariates$BrainRegion, paste(covariates$Diagnosis, covariates$Gender, sep = '.')))

# Get raw counts
counts = downloadFile('syn9765938') %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(counts) = counts$ensembl_gene_id
counts = counts[, covariates$SampleID]
all.used.ids = c(all.used.ids, 'syn9765629')

# Get gene specific parameters from synapse
gene.param = downloadFile('syn8449369')
all.used.ids = c(all.used.ids, 'syn8449369')

gene.len = dplyr::select(gene.param, ensembl_gene_id, gene.length) %>% 
  unique()
rownames(gene.len) = gene.len$ensembl_gene_id

gene.gc = dplyr::select(gene.param, ensembl_gene_id, percentage_gc_content) %>% 
  unique()
rownames(gene.gc) = gene.gc$ensembl_gene_id
```

### Filter genes
```{r filter.genes}
# Filter genes (remove genes that are expressed less than 1 cpm in 50% of samples per diagnosis per brain region)
lcpm = cpm(counts)

filtered.gene.ids = plyr::dlply(covariates, .(BrainRegion, Diagnosis), .fun = function(cov, logcpm){
  rownames(logcpm)[which(rowSums(logcpm[,cov$SampleID] >= 1)/dim(logcpm[,cov$SampleID])[2] >= 0.5)]
}, lcpm) %>%
  Reduce(c, .) %>%
  unique %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous")) %>%
  intersect(gene.len$ensembl_gene_id[!is.na(gene.len$gene.length)]) %>%
  intersect(gene.gc$ensembl_gene_id[!is.na(gene.gc$percentage_gc_content)])

counts = counts[filtered.gene.ids, ]
```
Choosing `r length(filtered.gene.ids)` genes expressed at 1 cpm in 50% of samples in at least one brain region and diagnosis category for further analysis

```{r fxn}
performDifferentialExpression <- function(cov, cnt, contVar, varOfInterest, grouping, gene.len, gene.gc){

  # Get design matrix
  designMat = getDesignMatrix(cov[,varOfInterest], Intercept = F)$design
  designMat = designMat[,linColumnFinder(designMat)$indepCols]

  # Compute offset for gene length and gc content
  cqn.gene.expression = cqn(cnt,
                            x = gene.gc[rownames(cnt), 'percentage_gc_content'],
                            lengths = gene.len[rownames(cnt), 'gene.length'],
                            lengthMethod = "smooth",
                            verbose = FALSE)
  cqn.gene.expression$E = cqn.gene.expression$y+cqn.gene.expression$offset

  # Identify outlier expression and assign NA values
  log.mat = apply(cqn.gene.expression$E, 1, function(x){
    mn = mean(x, na.rm = T)
    std.dev = sd(x, na.rm = T)
    return((x < (mn-3*std.dev)) | (x > (mn+3*std.dev)))
  }) %>% t
  cqn.gene.expression$E[log.mat] = NA

  # Estimate voom weights for dispersion control
  voom.weights = voom(cnt, design = designMat, plot=F, na.rm = T)

  # Estimate correlated effects
  correlation = parallelDuplicateCorrelation(voom.weights, design = designMat, block = grouping)

  # Fit linear model using new weights and new design
  baseModel = lmFit(cqn.gene.expression$E,
                    design = designMat,
                    weights = voom.weights$weights,
                    correlation = correlation$cor)

  # Perform differential expression
  contrasts = levels(cov[,contVar]) %>%
    paste(contVar,.,sep = '') %>%
    combn(2) %>%
    apply(2,paste, collapse = '-') %>%
    makeContrasts(contrasts = ., levels = colnames(baseModel$coefficients))
  baseModel.cntr =  contrasts.fit(baseModel, contrasts)
  baseModel.cntr = eBayes(baseModel.cntr)
  diffExp = lapply(1:dim(contrasts)[2], function(i, baseModel.cntr){
    topTable(baseModel.cntr, coef = i, number = 50000, confint = T) %>%
      rownameToFirstColumn('ensembl_gene_id')
    }, baseModel.cntr)
  names(diffExp) = colnames(contrasts)
  diffExp = diffExp %>%
    rbindlist(use.names = T, fill = T, idcol = 'Comparison') %>%
    dplyr::mutate(Comparison = gsub(contVar,'',Comparison))

  return(diffExp)
}
```

### Perform differential expression (per brain region)
```{r diff.exp}
all.results = plyr::dlply(covariates, .(Study), .fun = function(cov, cnt, gene.len, gene.gc){

  # Filter counts and covariates
  cnt = cnt[,cov$SampleID]
  grouping = cov$Donor_ID

  # Formulate covariates matrix
  cov$BrainRegion.Diagnosis = paste(cov$BrainRegion, cov$Diagnosis, sep = '.')
  FactorCov = c('BrainRegion.Diagnosis', 'Batch', 'Gender')
  ContCov = c('RIN', 'AAD', 'PMI',  'PCT_PF_READS_ALIGNED', 'PCT_CODING_BASES', 
              'PCT_INTERGENIC_BASES', 'PCT_INTRONIC_BASES', 'PCT_RIBOSOMAL_BASES')
  cov[,FactorCov] = lapply(cov[,FactorCov], as.factor)
  cov[,ContCov] = lapply(cov[,ContCov], as.numeric)
  rownames(cov) = cov$SampleID
  cov = cov[,c(FactorCov, ContCov)]
  # assocStat = getAssociationStatistics(cov)

  # Perform differential expression between case-control
  varOfInterest = c('BrainRegion.Diagnosis', 'Batch', 'Gender', 'RIN', 'AAD', 'PMI', 
                    'PCT_PF_READS_ALIGNED', 'PCT_INTERGENIC_BASES', 'PCT_INTRONIC_BASES', 
                    'PCT_RIBOSOMAL_BASES')
  contVar = 'BrainRegion.Diagnosis'

  baseDiffExp <- performDifferentialExpression(cov, cnt, contVar, varOfInterest, grouping, gene.len, gene.gc)

  # Perform differential expression between case-control in a gender specific manner
  cov$BrainRegion.Diagnosis.Gender = factor(paste(cov$BrainRegion, cov$Diagnosis, sep = '.') %>%
                                              paste(cov$Gender, sep = '.'))
  varOfInterest = c('BrainRegion.Diagnosis.Gender', 'Batch', 'RIN', 'AAD', 'PMI', 
                    'PCT_PF_READS_ALIGNED', 'PCT_INTERGENIC_BASES', 'PCT_INTRONIC_BASES', 
                    'PCT_RIBOSOMAL_BASES')
  contVar = 'BrainRegion.Diagnosis.Gender'

  genderDiffExp <- performDifferentialExpression(cov, cnt, contVar, varOfInterest, grouping, gene.len, gene.gc)

  return(rbindlist(list(baseDiffExp, genderDiffExp), use.names = T, fill = T))
}, counts, gene.len, gene.gc)

# Display results
writeLines('Number of differentially expressed genes between AD-CONTROL at an FDR of 0.05 and fold change of 1.2 are')
plyr::ldply(all.results, function(x){
  x %>%
    tidyr::separate(Comparison, c('from','to'), sep = '-') %>%
    tidyr::separate(from, c('from.BrainRegion', 'from.Diagnosis','from.a','from.Gender'), sep = '\\.') %>%
    tidyr::separate(to, c('to.BrainRegion', 'to.Diagnosis','to.a','to.Gender'), sep = '\\.') %>%
    dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), from.Diagnosis == 'AD', to.Diagnosis == 'CONTROL',
                  is.na(from.Gender), is.na(to.Gender), from.BrainRegion == to.BrainRegion) %>%
    tidyr::unite(Comparison, from.BrainRegion, from.Diagnosis, to.Diagnosis, sep = '.') %>%
    dplyr::group_by(Comparison) %>%
    dplyr::summarise(ncount = length(unique(ensembl_gene_id)))
}) %>% 
  kable(row.names=F)

writeLines('Number of differentially expressed genes between AD.FEMALE-CONTROL.FEMALE at an FDR of 0.05 and fold change of 1.2 are')
plyr::ldply(all.results, function(x){
  x %>%
    tidyr::separate(Comparison, c('from','to'), sep = '-') %>%
    tidyr::separate(from, c('from.BrainRegion', 'from.Diagnosis','from.a','from.Gender'), sep = '\\.') %>%
    tidyr::separate(to, c('to.BrainRegion', 'to.Diagnosis','to.a','to.Gender'), sep = '\\.') %>%
    dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), 
                  from.Diagnosis == 'AD', to.Diagnosis == 'CONTROL',
                  from.Gender == 'FEMALE', to.Gender == 'FEMALE', 
                  from.BrainRegion == to.BrainRegion) %>%
    tidyr::unite(Comparison, from.BrainRegion, from.Gender, from.Diagnosis, to.Diagnosis, sep = '.') %>%
    dplyr::group_by(Comparison) %>%
    dplyr::summarise(ncount = length(unique(ensembl_gene_id)))
}) %>% 
  kable(row.names=F)

writeLines('Number of differentially expressed genes between AD.MALE-CONTROL.MALE at an FDR of 0.05 and fold change of 1.2 are')
plyr::ldply(all.results, function(x){
  x %>%
    tidyr::separate(Comparison, c('from','to'), sep = '-') %>%
    tidyr::separate(from, c('from.BrainRegion', 'from.Diagnosis','from.a','from.Gender'), sep = '\\.') %>%
    tidyr::separate(to, c('to.BrainRegion', 'to.Diagnosis','to.a','to.Gender'), sep = '\\.') %>%
    dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), 
                  from.Diagnosis == 'AD', to.Diagnosis == 'CONTROL',
                  from.Gender == 'MALE', to.Gender == 'MALE', 
                  from.BrainRegion == to.BrainRegion) %>%
    tidyr::unite(Comparison, from.BrainRegion, from.Gender, from.Diagnosis, to.Diagnosis, sep = '.') %>%
    dplyr::group_by(Comparison) %>%
    dplyr::summarise(ncount = length(unique(ensembl_gene_id)))
}) %>% 
  kable(row.names=F)

writeLines('Number of differentially expressed genes between AD.FEMALE-AD.MALE at an FDR of 0.05 and fold change of 1.2 are')
plyr::ldply(all.results, function(x){
  x %>%
    tidyr::separate(Comparison, c('from','to'), sep = '-') %>%
    tidyr::separate(from, c('from.BrainRegion', 'from.Diagnosis','from.a','from.Gender'), sep = '\\.') %>%
    tidyr::separate(to, c('to.BrainRegion', 'to.Diagnosis','to.a','to.Gender'), sep = '\\.') %>%
    dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), 
                  from.Diagnosis == 'AD', to.Diagnosis == 'AD',
                  from.Gender == 'FEMALE', to.Gender == 'MALE', 
                  from.BrainRegion == to.BrainRegion) %>%
    tidyr::unite(Comparison, from.BrainRegion, from.Diagnosis, from.Gender, to.Gender, sep = '.') %>%
    dplyr::group_by(Comparison) %>%
    dplyr::summarise(ncount = length(unique(ensembl_gene_id)))
}) %>% 
  kable(row.names=F)

writeLines('Number of differentially expressed genes between CONTROL.FEMALE-CONTROL.MALE at an FDR of 0.05 and fold change of 1.2 are')
plyr::ldply(all.results, function(x){
  x %>%
    tidyr::separate(Comparison, c('from','to'), sep = '-') %>%
    tidyr::separate(from, c('from.BrainRegion', 'from.Diagnosis','from.a','from.Gender'), sep = '\\.') %>%
    tidyr::separate(to, c('to.BrainRegion', 'to.Diagnosis','to.a','to.Gender'), sep = '\\.') %>%
    dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), 
                  from.Diagnosis == 'CONTROL', to.Diagnosis == 'CONTROL',
                  from.Gender == 'FEMALE', to.Gender == 'MALE', 
                  from.BrainRegion == to.BrainRegion) %>%
    tidyr::unite(Comparison, from.BrainRegion, from.Diagnosis, from.Gender, to.Gender, sep = '.') %>%
    dplyr::group_by(Comparison) %>%
    dplyr::summarise(ncount = length(unique(ensembl_gene_id)))
}) %>% 
  kable(row.names=F)
```

### Store results in synapse
```{r synapse.store, include=FALSE, cache=FALSE}
# Store consolidated data in synapse
CODE = Folder(name = 'Differential Expression Analysis (per study)', parentId = parentId)
CODE = synStore(CODE)

# Store results in synapse
save(list = 'all.results', file = 'AllDifferentialExpressionResults.RData')
OBJ = File('AllDifferentialExpressionResults.RData', 
           name = 'All Differential Expression Results', 
           parentId = CODE$properties$id)
OBJ = synStore(OBJ,
               used = all.used.ids,
               executed = thisFile,
               activityName = activityName,
               activityDescription = activityDescription)
```
### Source code in github
[Source Code](`r thisFile`)