### 2017-07-07 code/Rmd/ALL_REPROCESSED.Rmd
Minor bug fix

---
title: "Differential expression analysis of reprocessed AMPAD RNASeq data (All 7 brain regions with CQN normlaisation and BIC based model selection)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./ALL_REPROCESSED.Rmd",
                                 parentId = "syn8672415",
                                 entityName = 'ALL Reprocessed RNASeq (Normalised)')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(limma)
library(psych)
library(edgeR)
library(cqn)
library(lme4)
library(biomaRt)

library(synapseClient)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synapseLogin()

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn8672415';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of reprocessed rnaseq from all 7 brain regions';

thisFileName <- 'ALL_REPROCESSED.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_reprocessing')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Data download
```{r download.data, cache=TRUE}
# Get covariates from synapse
covariates.ids = c(ROSMAP = 'syn8456631',
                   MSBB = 'syn8484996',
                   MAYO = 'syn8466814')
all.used.ids = as.character(covariates.ids)
covariates = lapply(covariates.ids, function(id){
  read.table(synGet(id)@filePath, header = T, sep = '\t')
})

covariates$ROSMAP = covariates$ROSMAP %>%
  dplyr::mutate(SampleID = as.character(SampleID),
                BrainRegion = 'DLPFC',
                BrainRegion.Diagnosis = factor(paste(BrainRegion, Diagnosis, sep = '.')),
                msex = factor(msex, levels = c('1', '0'), labels = c('1'='MALE','0'='FEMALE')),
                Batch = factor(Batch)) %>%
  dplyr::rename(Gender = msex, RIN = RINcontinuous, AOD = age_death, PMI = pmi) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis, Batch, Gender, AOD, PMI, RIN, PCT_PF_READS_ALIGNED, 
                PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)
covariates$ROSMAP$individualIdentifier = factor(paste0('ROSMAPpseudoid',1:dim(covariates$ROSMAP)[1]))

covariates$MSBB = covariates$MSBB %>%
  dplyr::mutate(SampleID = as.character(SampleID),
                individualIdentifier = factor(paste0('BB',individualIdentifier)),
                SEX = factor(SEX, levels = c('M', 'F'), labels = c('M'='MALE','F'='FEMALE')),
                batch = factor(batch),
                PMI = PMI/60) %>%
  dplyr::rename(Batch = batch, Gender = SEX) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis, individualIdentifier, Batch, Gender, AOD, PMI, RIN, 
                PCT_PF_READS_ALIGNED, PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)

covariates$MAYO = covariates$MAYO %>%
  dplyr::mutate(SampleID = as.character(SampleID),
                Donor_ID = factor(paste0('MAYO', Donor_ID)),
                Gender = factor(Gender, levels = c('M', 'F'), labels = c('M'='MALE','F'='FEMALE')),
                Batch = factor(Source)) %>%
  dplyr::rename(individualIdentifier = Donor_ID, AOD = AgeAtDeath) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis, individualIdentifier, Batch, Gender, AOD, PMI, RIN, 
                PCT_PF_READS_ALIGNED, PCT_CODING_BASES, PCT_INTERGENIC_BASES, PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES)

# Download sample counts
counts.id = c(ROSMAP = 'syn8691134',
              MSBB = 'syn8691099',
              MAYO1 = 'syn8690799', 
              MAYO2 = 'syn8690904')
all.used.ids = c(all.used.ids, as.character(counts.id))
counts = lapply(counts.id, function(id){
  tmp = read.table(synGet(id)@filePath, header=T, sep='\t', check.names = F)
}) %>%
  plyr::join_all(type = 'full')
```

```{r preprocess.data1, results='asis'}
# Filter OTHER samples 
covariates = rbindlist(covariates, use.names = T, fill = T, idcol = 'Study') %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::filter(Diagnosis != 'OTHER') %>%
  dplyr::mutate(BrainRegion = factor(BrainRegion, levels = c('DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'CER', 'TCX')),
                BrainRegion.Diagnosis = factor(paste(BrainRegion, Diagnosis, sep = '.')))

# Match covariates to expression data
indToRetain = intersect(covariates$SampleID, colnames(counts))

rownames(counts) = counts$feature
counts = counts[,indToRetain]

rownames(covariates) = covariates$SampleID
covariates = covariates[indToRetain,]

covariates %>% 
  dplyr::group_by(BrainRegion, Diagnosis) %>% 
  dplyr::summarise(count = length(unique(SampleID))) %>% 
  tidyr::spread(BrainRegion, count) %>%
  kable()

COUNT = counts
```

```{r gene.param}
# Get gene specific parameters from synapse
GENE.PARAM = downloadFile('syn8449369')
all.used.ids = c(all.used.ids, 'syn8449369')

GENE.LEN = dplyr::select(GENE.PARAM, Gene.ID, gene.length) %>% unique()
rownames(GENE.LEN) = GENE.LEN$Gene.ID

GENE.GC = dplyr::select(GENE.PARAM, Gene.ID, percentage_gc_content) %>% unique()
rownames(GENE.GC) = GENE.GC$Gene.ID

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "gene_biotype"),
                       filters = "ensembl_gene_id", values = GENE.PARAM$ensembl_gene_id,
                       mart = mart)
GENE.PARAM = left_join(GENE.PARAM, Ensemble2HGNC)
```

### Covariates clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c("individualIdentifier", "Study", "BrainRegion.Diagnosis", "BrainRegion", "Diagnosis",
                      "Gender", "Batch")
ContCovariates <- c("AOD", "PMI", "RIN", "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES", "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES",
                    "PCT_RIBOSOMAL_BASES")

# Find inter relation between factor covariates
COVARIATES = covariates[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- covariates$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})
COVARIATES$BrainRegion = factor(COVARIATES$BrainRegion, levels = c('DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'CER', 'TCX'))

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, name = 'EffectSize',
            cluster_rows = T, cluster_columns = T, 
            col = colorRamp2(c(-1,0,1),c('blue','white','red')),
            cell_fun = function(j,i,x,y,wifht,height,fill){
              grid.text(COVARIATES.CORRELATION$PVAL[i,j] %>% format(digits = 1), x, y)
            })
draw(h, heatmap_legend_side = 'left', padding = unit(c(20,2,2,15), 'mm'))
```

## Explore metatdata
```{r data.explore, fig.width = 16, fig.height = 8}
my.theme <- theme(legend.position = 'top', axis.text.x = element_text(angle = 45, hjust = 1), plot.title=element_text(hjust=0.5))

# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = BrainRegion, y = RIN, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme

# Age of Death
p[[2]] = ggplot(COVARIATES, aes(x = BrainRegion, y = AOD, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('Age Of Death') + my.theme

# PMI
p[[3]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PMI, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme

# Intronic Rate
p[[4]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PCT_INTRONIC_BASES, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Intronic Fractions') + my.theme

# IntergenicRate
p[[5]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PCT_INTERGENIC_BASES, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Intergenic Fractions') + my.theme

# Ribosomal Fractions
p[[6]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PCT_RIBOSOMAL_BASES, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Ribosomal Fractions') + my.theme

# Reads Aligned
p[[7]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PCT_PF_READS_ALIGNED, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[7]] = p[[7]] + ggtitle('Fraction Reads Aligned') + my.theme

# Coding Bases
p[[8]] = ggplot(COVARIATES, aes(x = BrainRegion, y = PCT_CODING_BASES, color = Study, fill = Diagnosis)) + geom_boxplot()
p[[8]] = p[[8]] + ggtitle('Fraction Coding Bases') + my.theme

multiplot(plotlist = p, cols = 4)
```
```{r data.explore1, fig.width = 6, fig.height = 6}
# BrainRegion
vcd::mosaic(~ BrainRegion + Diagnosis, data = COVARIATES)

# Gender
vcd::mosaic(~ BrainRegion + Gender, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per BrainRegion.Diagnosis. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
genesToAnalyze = plyr::dlply(covariates, .(BrainRegion.Diagnosis), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$Gene.ID[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$Gene.ID[!is.na(GENE.LEN$gene.length)])

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(Gene.ID = genes) %>%
  dplyr::left_join(GENE.PARAM) %>%
  dplyr::group_by(gene_biotype) %>%
  dplyr::summarise(fraction  = length(unique(Gene.ID))) %>%
  dplyr::filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis.

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
indToRemove = c('367_120502')

# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID)
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''

sdev = round((PC$sdev^2/sum(PC$sdev^2)) * 100, 0)

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") + ylab(paste0('PC1(',sdev[2],'%)')) + xlab(paste0('PC1(',sdev[1],'%)'))
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(BrainRegion.Diagnosis, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(BrainRegion+.~Diagnosis, scale = 'free')
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dplyr::group_by(BrainRegion, Diagnosis) %>%
  dplyr::summarise(count = length(unique(SampleID))) %>%
  tidyr::spread(BrainRegion, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$individualIdentifier))` unique individuals

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are:
`r kable(tmp)`

### Winsorise counts (Gene outliers in samples)
```{r winsorise.data}
LOG.CPM = cpm(PROCESSED_COUNTS$filteredExprMatrix$counts, log = T)

# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(LOG.CPM, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))

  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t

# Back calculate counts for further analysis
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(NEW.COUNTS, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset

# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))

  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
```

### Sample clustering
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="right")
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("BrainRegion", "Gender", "Diagnosis", "Study")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```
```{r temp, include = F}
dev.off()
gc()
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE, fig.height=10, fig.width=20}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$E) %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(BrainRegion.Diagnosis, c('Diagnosis', 'BrainRegion'), sep = '_'))

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(BrainRegion+.~Diagnosis, scale = 'free')
p
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, cache=TRUE}
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preAdjustedSigCovars.NULL, fig.width=20, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Model Identification
```{r fxn}
# Function to fit a mixed effect model and return BIC
fitMixedEffectModel <- function(varsToFit, groupingVar, lcpm, cnt, covars){

  # Get design matrix
  designMat = getDesignMatrix(covars[colnames(lcpm), varsToFit, drop=F], Intercept = F)
  designMat = designMat$design[,linColumnFinder(designMat$design)$indepCols]

  # Estimate voom weights
  voomWeights = limma::voom(cnt[rownames(lcpm),colnames(lcpm)], design = designMat, plot=F)$weights

  modelBIC = foreach::foreach(i = 1:dim(lcpm)[1],
                              .combine = rbind,
                              .packages = c('lme4', 'dplyr'),
                              .export = c('voomWeights')) %dopar% {

                                # Perform weighted mixed linear models for every gene
                                expr = data.frame(Gx = lcpm[i,])
                                data = cbind(expr,
                                             designMat[rownames(expr),],
                                             covars[rownames(designMat),groupingVar,drop=F])
                                formula = paste(setdiff(colnames(data), c('Gx', groupingVar)), collapse = '+') %>%
                                  paste('Gx ~',.) %>%
                                  c(.,paste0('(1|',groupingVar,')')) %>%
                                  paste(collapse = '+')

                                weights = voomWeights[i,]
                                BIC = NA
                                tryCatch({
                                  mdl = lmer(formula = formula, 
                                             data = data, 
                                             weights = weights)
                                  BIC = BIC(mdl)
                                }, error = function(e){
                                  BIC = NA
                                })
                              return(data.frame(Gene.ID = rownames(lcpm)[i],
                                                BIC = BIC))
                            }
  return(modelBIC)
  }

# Function to select covariates in a step-wise regression fashion
selectCovarsToAdjust <- function(covarsToSelect, adjustCovars, covars, lcpm, cnt, baseModelBIC){

  varsSelected = c()
  pkgs = c('CovariateAnalysis', 'data.table', 'dplyr', 'plyr',
           'edgeR', 'limma', 'lme4', 'foreach', 'doParallel')
  while(length(covarsToSelect) != 0){
    allModelBIC = foreach(i = 1:length(covarsToSelect),
                          .combine = c,
                          .packages = pkgs,
                          .export = c('fitMixedEffectModel')) %dopar% {

                            # Fit a mixed effect linear model and get all BIC
                            modelBIC = fitMixedEffectModel(c(adjustCovars, covarsToSelect[i]),
                                                           c('individualIdentifier'), 
                                                           lcpm, 
                                                           cnt, 
                                                           covars) %>%
                            plyr::rename(c('BIC' = covarsToSelect[i]))

                            return(list(modelBIC))
                          }
    allModelBIC = plyr::join_all(allModelBIC)

    # Find number of genes that are changed based on BIC
    nChange = allModelBIC %>%
      tidyr::gather(ModelName, Value, -Gene.ID) %>%
      dplyr::left_join(baseModelBIC) %>%
      dplyr::mutate(diff = baseModel - Value) %>%
      dplyr::group_by(ModelName) %>%
      dplyr::summarise(better = sum(diff >= 5), 
                       worse = sum(diff <= -5)) %>%
      dplyr::mutate(nchange = (better/(better+worse)) * 100)

    print(nChange)

    # Select variable
    tmp = nChange %>%
      dplyr::filter(nchange > 1) %>%
      dplyr::top_n(1, nchange)

    if (dim(tmp)[1] == 0){
      covarsToSelect = c()
    } else {
      baseModelBIC = allModelBIC %>%
        dplyr::select(Gene.ID, one_of(tmp$ModelName))
      colnames(baseModelBIC)[2] = 'baseModel'
      varsSelected = c(varsSelected, tmp$ModelName)
      covarsToSelect = setdiff(covarsToSelect, tmp$ModelName)
      print(paste('Variable chosen to be added:',tmp$ModelName))
    }
  }

  return(varsSelected)
}
```
#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1}
# Fit a base model with the following covariates
adjust.covars = c('BrainRegion.Diagnosis')
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   CQN.GENE_EXPRESSION$E, 
                                   NEW.COUNTS, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

# Iteratively identify covariates in phase I
phase1.covars = c('Study', 'Gender', 'PMI', 'AOD', 'RIN', 'RIN2')
vars.selected = selectCovarsToAdjust(phase1.covars, adjust.covars, 
                                     COVARIATES, CQN.GENE_EXPRESSION$E, 
                                     NEW.COUNTS, baseModelBIC)

adjust.covars = c(adjust.covars, vars.selected)
```
Clinical and technical variables selected in phase I are `r paste(adjust.covars, collapse = ', ')`

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   CQN.GENE_EXPRESSION$E, 
                                   NEW.COUNTS, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

# Iteratively identify covariates in phase I
phase2.covars = c('Batch')
vars.selected = selectCovarsToAdjust(phase2.covars, adjust.covars, 
                                     COVARIATES, CQN.GENE_EXPRESSION$E, 
                                     NEW.COUNTS, baseModelBIC)

adjust.covars = c(adjust.covars, vars.selected)
```
Clinical, technical and batch variables selected after phase I and II of model identification are `r paste(adjust.covars, collapse = ', ')`

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   CQN.GENE_EXPRESSION$E, 
                                   NEW.COUNTS, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

# Iteratively identify covariates in phase I
phase3.covars = c("PCT_PF_READS_ALIGNED", "PCT_CODING_BASES", "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES")
vars.selected = selectCovarsToAdjust(phase3.covars, adjust.covars, 
                                     COVARIATES, CQN.GENE_EXPRESSION$E, 
                                     NEW.COUNTS, baseModelBIC)

adjust.covars = c(adjust.covars, vars.selected)
```
Clinical, technical and batch variables selected after phase I, II and III of model identification are `r paste(adjust.covars, collapse = ', ')`

### Normalisation
1. BrainRegion.Diagnosis is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis and tissue)
2. Individual_ID is chosen as random effect
```{r iterative.normalisation, results='asis'}
writeLines(paste('Using following covariates in the model:',
                 paste(adjust.covars, collapse=', '),
                 'as fixed effects and Individual_ID is chosen as random effect'))

# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,adjust.covars,drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

# Estimate voom weights
VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, design=DM1$design, plot=F)

# Estimate consensus correlation coefficient
correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION, block = COVARIATES$Individual_ID)

# Re-calculate voom weights with correlation of random effects
VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, 
                            design=DM1$design, plot=F,
                            block = COVARIATES$Individual_ID, 
                            correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, CQN.GENE_EXPRESSION$E)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
```

### Sanity check
```{r residualSigCovars.manual, fig.width=12, fig.height=8}
residualSigCovars[["PC_res"]][[2]]$plotData
```

```{r decompse.RESIDUALS, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="right")
p
```

### Residual calculation
Calculate weighted residuals and add back "BrainRegion.Diagnosis" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("BrainRegion.Diagnosis", colnames(DM1$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
  ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

### Coexpression of residual expression 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="right")
p
```

```{r decompse.normalise.data2.1, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('BrainRegion', 'Diagnosis', 'Gender', 'Study')])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

```{r temp1, include=F}
dev.off()
gc()
```

### Differential expression analysis (with BrainRegion.Diagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp1, fig.height=7, fig.width=24}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, adjust.covars, drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, design=DESIGN$design, plot=F)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

# Re-estimate voom weights
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  tmp = combn(paste0('BrainRegion.Diagnosis', region, '_', c('AD', 'CONTROL')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}

contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT.CONTR){
  topTable(FIT.CONTR, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Gene.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.Diagnosis','',Comparison),
                Study = 'ALL',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  left_join(GENE.PARAM %>%
              dplyr::select(Gene.ID, ensembl_gene_id, hgnc_symbol, percentage_gc_content, gene.length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position='top')
p = p + facet_grid(.~Comparison, scales = 'fixed')
p

all.diff.exp = list(Diagnosis = DE)
all.fit = list(Diagnosis = FIT)
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=20, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE, aes(x = log10(gene.length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(gene.length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE, aes(x = percentage_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE, aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (with BrainRegion.Diagnosis and Gender)
Differential expression is performed on the BrainRegion.Diagnosis x Gender variable by controlling for covariates identified above
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp2, fig.height=7, fig.width=24}
## Modeling Diagnosis, msex and age_death conjointly
COVARIATES$BrainRegion.Diagnosis.Gender = factor(paste(COVARIATES$BrainRegion.Diagnosis,
                                                       COVARIATES$Gender,
                                                       sep = '_'))

# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('BrainRegion.Diagnosis.Gender', setdiff(adjust.covars, c('BrainRegion.Diagnosis', 'Gender'))), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, design=DESIGN$design, plot=F)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

# Re-estimate voom weights
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0('BrainRegion.Diagnosis.Gender', region, '_AD_MALE-BrainRegion.Diagnosis.Gender', region, '_CONTROL_MALE'),
          paste0('BrainRegion.Diagnosis.Gender', region, '_AD_FEMALE-BrainRegion.Diagnosis.Gender', region, '_CONTROL_FEMALE'),
          paste0('0.5*BrainRegion.Diagnosis.Gender', region, '_AD_MALE+0.5*BrainRegion.Diagnosis.Gender', region, 
                 '_AD_FEMALE-0.5*BrainRegion.Diagnosis.Gender', region, '_CONTROL_MALE-0.5*BrainRegion.Diagnosis.Gender',region, '_CONTROL_FEMALE'),
          paste0('0.5*BrainRegion.Diagnosis.Gender', region, '_AD_FEMALE+0.5*BrainRegion.Diagnosis.Gender', region, 
                 '_CONTROL_FEMALE-0.5*BrainRegion.Diagnosis.Gender', region, '_AD_MALE-0.5*BrainRegion.Diagnosis.Gender',region, '_CONTROL_MALE'),
          paste0('0.5*BrainRegion.Diagnosis.Gender', region, '_AD_FEMALE-0.5*BrainRegion.Diagnosis.Gender', region, 
                 '_CONTROL_FEMALE-0.5*BrainRegion.Diagnosis.Gender', region, '_AD_MALE+0.5*BrainRegion.Diagnosis.Gender',region, '_CONTROL_MALE'))
  cntr = c(cntr, cnt)
}
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
cnames = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0(region, '_AD_MALE-', region, '_CONTROL_MALE'),
          paste0(region, '_AD_FEMALE-', region, '_CONTROL_FEMALE'),
          paste0(region, '_AD_ALL-', region, '_CONTROL_ALL'),
          paste0(region, '_ALL_FEMALE-', region, '_ALL_MALE'),
          paste0(region, '_ALL_ALL-', region,'_ALL_ALL'))
  cnames = c(cnames, cnt)
}
colnames(contrast) = cnames

FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Gene.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  left_join(GENE.PARAM %>%
              dplyr::select(Gene.ID, hgnc_symbol, percentage_gc_content, gene.length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Gene.ID, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Gene.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

all.diff.exp = c(all.diff.exp, list(Diagnosis.Gender = DE))
all.fit = c(all.fit, list(Diagnosis.Gender = FIT))
```

### Differential expression analysis (with BrainRegion.Diagnosis and AOD)
Differential expression is performed on the BrainRegion.Diagnosis x AOD variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp3, fig.height=7, fig.width=24}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(setdiff(adjust.covars, 'AOD')), drop = F], Intercept = F)
ind = grep('BrainRegion.Diagnosis', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind]*COVARIATES$AOD
colnames(DESIGN$design)[ind] = paste(colnames(DESIGN$design)[ind], 'AOD', sep = '_')
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, design=DESIGN$design, plot=F)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

# Re-estimate voom weights
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0('BrainRegion.Diagnosis', region, '_AD_AOD-BrainRegion.Diagnosis', region, '_CONTROL_AOD'),
          paste0('0.5*BrainRegion.Diagnosis', region, '_AD_AOD+0.5*BrainRegion.Diagnosis', region, '_CONTROL_AOD'))
  cntr = c(cntr, cnt)
}
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
cnames = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0(region, '_AD_AOD-', region, '_CONTROL_AOD'),
          paste0(region, '_ALL_AOD'))
  cnames = c(cnames, cnt)
}
colnames(contrast) = cnames

FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Gene.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  left_join(GENE.PARAM %>%
              dplyr::select(Gene.ID, hgnc_symbol, percentage_gc_content, gene.length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Gene.ID, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Gene.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
```

### Differential expression analysis (with BrainRegion.Diagnosis and Celltype Fractions)
Differential expression is performed on the BrainRegion.Diagnosis x CelltypeFractions variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp4.1, fig.height=10, fig.width=15}
# Cell type markers
CELL.FRAC = downloadFile('syn10163089')
all.used.ids = c(all.used.ids, 'syn10163089')

# Merge covariates to cell type fraction
NEW.COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID') %>%
  left_join(CELL.FRAC)

my.theme = theme(legend.position = 'top', axis.text.x = element_text(angle = 45, hjust = 1))
pl = list()
pl[[1]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = astrocytes, color = Diagnosis)) + geom_boxplot() + ggtitle('Astrocytes') + my.theme
pl[[2]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = endothelial, color = Diagnosis)) + geom_boxplot() + ggtitle('Endothelial') + my.theme
pl[[3]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = microglia, color = Diagnosis)) + geom_boxplot() + ggtitle('Microglia') + my.theme
pl[[4]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = neurons, color = Diagnosis)) + geom_boxplot() + ggtitle('Neurons') + my.theme
pl[[5]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = myeloid, color = Diagnosis)) + geom_boxplot() + ggtitle('Myeloid') + my.theme
pl[[6]] = ggplot(NEW.COVARIATES, aes(x = BrainRegion, y = oligodendrocytes, color = Diagnosis)) + geom_boxplot() + ggtitle('Oligodendrocytes') + my.theme
multiplot(plotlist = pl, cols = 3)

# Get design matrix
DESIGN = getDesignMatrix(NEW.COVARIATES[, c(adjust.covars, "astrocytes", "endothelial", "microglia", "myeloid", "neurons", "oligodendrocytes", "OPC"), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, design=DESIGN$design, plot=F)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

# Re-estimate voom weights
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0('BrainRegion.Diagnosis', region, '_AD-BrainRegion.Diagnosis', region, '_CONTROL'))
  cntr = c(cntr, cnt)
}
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
cnames = c()
for (region in  c('DLPFC', 'FP', 'STG', 'PHG', 'IFG', 'CER', 'TCX')){
  cnt = c(paste0(region, '_AD_CELLFRAC-', region, '_CONTROL_CELLFRAC'))
  cnames = c(cnames, cnt)
}
colnames(contrast) = cnames

FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Gene.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  left_join(GENE.PARAM %>%
              dplyr::select(Gene.ID, hgnc_symbol, percentage_gc_content, gene.length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Gene.ID, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Gene.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

all.diff.exp = c(all.diff.exp, list(Diagnosis.CellFraction = DE))
all.fit = c(all.fit, list(Diagnosis.CellFraction = FIT))
```

### Gene set enrichment analysis
Using Fisher's exact test for gene set enrichment analysis
```{r enrich.de, fig.height=16, fig.width=16}
# Perform gene set enrichment analysis using wilcox test from limma
diffexp.genes = all.diff.exp %>%
  data.table::rbindlist(idcol='Model', use.names = T, fill = T) %>%
  dplyr::mutate(Direction = sign(logFC),
                Direction = factor(Direction, labels = c('-1' = 'DOWN', '1' = 'UP')))

geneSet.test = diffexp.genes %>%
  plyr::dlply(.(Model, Comparison, Direction), .fun = function(x){
    unique(x$hgnc_symbol[x$adj.P.Val <= 0.05])
  })

statistics.test = diffexp.genes %>%
  plyr::dlply(.(Model, Comparison), .fun = function(x){
    tmp = dplyr::select(x, hgnc_symbol, adj.P.Val) %>% dplyr::group_by(hgnc_symbol) %>% dplyr::top_n(1,-adj.P.Val) %>% unique() %>% as.data.frame()
    rownames(tmp) = tmp$hgnc_symbol
    return(tmp)
  })
names(statistics.test) = paste(names(statistics.test),'UP',sep = '.')
statistics.test1 = statistics.test
names(statistics.test1) = gsub('.UP','.DOWN', names(statistics.test1))
statistics.test = c(statistics.test, statistics.test1)

backgroundGenes = unique(DE$hgnc_symbol)

# Download all related genesets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c(all.used.ids, 'syn4867851')
geneSet.ref = GeneSets[c('KEGG_2016', 'GO_Biological_Process_2015', 'Reactome_2016', 'WikiPathways_2016')] %>%
  CovariateAnalysis::filterGeneSets(backgroundGenes) %>%
  do.call(c,.)

# Perform enrichment analysis
enrichment.results = mapply(function(gs.test, st.test, gs.ref){
  plyr::ldply(gs.ref, .fun = function(y, gs.test, st.test){
    tmp = CovariateAnalysis::fisherEnrichment(gs.test, y, st.test$hgnc_symbol)
    tmp$wilcox.pval = limma::geneSetTest(which(st.test$hgnc_symbol %in% y), st.test[,'adj.P.Val'])
    return(tmp)
  }, gs.test, st.test, .parallel = T, .id = 'GeneSetName') %>%
    dplyr::mutate(fdr = p.adjust(pval, method = 'fdr'),
                  wilcox.fdr = p.adjust(wilcox.pval, method = 'fdr'))
}, geneSet.test, statistics.test[names(geneSet.test)], 
MoreArgs = list(geneSet.ref),
SIMPLIFY = F) %>%
  data.table::rbindlist(idcol = 'Comparison', use.names = T, fill = T)
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  assaySubType = 'Reprocessed',

  tissueTypeAbrv	= 'DLPFC, FP, STG, PHG, IFG, CER, TCX', 
  study = 'ALL', 
  center = 'ALL',

  organism = 'HomoSapiens',
  consortium	= 'AMP-AD',

  normalizationStatus	= TRUE,
  modelSystem	= FALSE,

  normalizationType	= 'CQN'
)

# Function to change entity ACL (also creates new acls if it doesn't exist)
synChangeEntityACL <- function(id, new.racl){
  tryCatch({ 
    # If old acl exist update
    old.acl = synGetEntityACL(id)
    old.acl@resourceAccess = new.racl
    synUpdateEntityACL(old.acl)
  }, error = function(e){ 
    # If not create a new acl
    new.acl = AccessControlList(id = id, resourceAccess = new.racl)
    synCreateEntityACL(new.acl)
  })
}

# Create new resource acls
new.racl = ResourceAccess(principalId = '3319864', 
                          accessType = c('UPDATE', 'CHANGE_SETTINGS', 'DELETE', 'READ', 
                                         'DOWNLOAD', 'MODERATE', 'CREATE', 'CHANGE_PERMISSIONS')) %>%
  ResourceAccessList()

# Code
CODE <- Folder(name = "ALL Reprocessed RNASeq (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
synChangeEntityACL(CODE@properties$id, new.racl)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'AMP_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('AMP_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = all.used.ids, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
synChangeEntityACL(COV_OBJ@properties$id, new.racl)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('Gene.ID') %>%
  write.table(file = 'AMP_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('AMP_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  write.table(file = 'AMP_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('AMP_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('Gene.ID') %>%
  write.table(file = 'AMP_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('AMP_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DESIGN$design %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'AMP_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('AMP_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
synChangeEntityACL(DM_OBJ@properties$id, new.racl)

# Store differential expression results
write.table(diffexp.genes, file = 'AMP_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('AMP_DiffExpression.tsv', 
                name = 'Differential Expression Results', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store enrichment results
write.table(enrichment.results, file = 'AMP_EnrichmentResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('AMP_EnrichmentResults.tsv', 
                  name = 'GeneSet Enrichment Results', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichment'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store genesets
save(list = 'geneSet.test', file = 'AMP_DiffExpGeneSet.RData')
DEG_OBJ = File('AMP_DiffExpGeneSet.RData',
               name = 'Differential Expression Gene Set', 
               parentId = CODE$properties$id)
annotations(DEG_OBJ) = annotations(CODE)
annotations(DEG_OBJ)$dataSubType = 'diffExpGS'
DEG_OBJ = synStore(DEG_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

synDeleteEntityACL(CODE@properties$id)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ----------------------------   |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Results                         |  `r ENRICH_OBJ$properties$id`    |
|  Differentially Expressed Gene Sets         |  `r DEG_OBJ$properties$id`       |