### 2017-08-30 code/Rmd/CMC_HBCC_COMPARISON.Rmd

---
title: "Compare CMC and HBCC differential expression"
author: "Thanneer M Perumal"
date: "`r base::date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE, include=FALSE, echo=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::storeAndKnitToFileEntity(file = "./CMC_HBCC_COMPARISON.Rmd",
                                       parentId = "syn7501880",
                                       entityName = 'CMC - HBCC - Reproducibility Analysis')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(limma)
library(psych)
library(edgeR)
library(cqn)
library(biomaRt)

library(synapseClient)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synapseLogin()

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE, echo=FALSE}
parentId = 'syn8672415';
activityName = 'Reproducibility Analysis b/w cmc and hbcc samples';
activityDescription = 'Reproducibility Analysis b/w cmc and hbcc samples';

thisFileName <- 'CMC_HBCC_COMPARISON.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='CMC')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Data download
```{r download.data, cache=TRUE}
# Get differential expression results
diff.exp.ids = c(CMC = 'syn9874406',
                 CMC_qSVA = 'syn10341319',
                 HBCC = 'syn10263511',
                 HBCC_qSVA = 'syn10294725',
                 MATCHED = 'syn10490519')

diff.exp = lapply(diff.exp.ids, downloadFile) 
diff.exp$HBCC = diff.exp$HBCC %>%
  tidyr::unite(a, to.Dx, to.Tissue, sep = '_') %>%
  tidyr::unite(b, ref.Dx, ref.Tissue, sep = '_') %>%
  tidyr::unite(Comparison, a, b, sep = '-')
diff.exp$HBCC_qSVA = diff.exp$HBCC_qSVA %>%
  tidyr::unite(a, to.Dx, to.Tissue, sep = '_') %>%
  tidyr::unite(b, ref.Dx, ref.Tissue, sep = '_') %>%
  tidyr::unite(Comparison, a, b, sep = '-')

diff.exp = data.table::rbindlist(diff.exp, use.names = T, fill = T, idcol = 'Model')
```

### Analyse results
Replication rate based on number of gene differentiallty expressed in the same directiona at an FDR of 0.05
```{r anlz.results}
rep.rate = list()
for (pval in seq(0,-4,-1)){
  de.genes = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC') %>%
    dplyr::mutate(Direction = factor(sign(logFC), levels = c('-1','1'), labels = c('-1' = 'DOWN','1' = 'UP'))) %>%
    plyr::dlply(.(Model, Direction), .fun = function(x){
      x$ensembl_gene_id[x$adj.P.Val <= 10^pval] %>% unique()
    })

  l1 = length(intersect(de.genes$CMC.DOWN, de.genes$HBCC.DOWN))
  l2 = length(intersect(de.genes$CMC.UP, de.genes$HBCC.UP))
  l3 = length(unique(c(de.genes$CMC.UP, de.genes$HBCC.UP, de.genes$CMC.DOWN, de.genes$HBCC.DOWN)))
  wo.qsva = (l1+l2)/l3
  n.wo.qsva = l1+l2

  l1 = length(intersect(de.genes$CMC_qSVA.DOWN, de.genes$HBCC_qSVA.DOWN))
  l2 = length(intersect(de.genes$CMC_qSVA.UP, de.genes$HBCC_qSVA.UP))
  l3 = length(unique(c(de.genes$CMC_qSVA.UP, de.genes$HBCC_qSVA.UP, de.genes$CMC_qSVA.DOWN, de.genes$HBCC_qSVA.DOWN)))
  w.qsva = (l1+l2)/l3
  n.w.qsva = l1+l2

  de.genes = diff.exp %>%
    dplyr::filter(Comparison %in% c('SCZ_DLPFC_CMC-Control_DLPFC_CMC', 'SCZ_DLPFC_HBCC-Control_DLPFC_HBCC')) %>%
    dplyr::mutate(Direction = factor(sign(logFC), levels = c('-1','1'), labels = c('-1' = 'DOWN','1' = 'UP'))) %>%
    plyr::dlply(.(Comparison, Model, Direction), .fun = function(x){
      x$ensembl_gene_id[x$adj.P.Val <= 10^pval] %>% unique()
    })

  l1 = length(intersect(de.genes$`SCZ_DLPFC_CMC-Control_DLPFC_CMC.MATCHED.DOWN`,
                        de.genes$`SCZ_DLPFC_HBCC-Control_DLPFC_HBCC.MATCHED.DOWN`))
  l2 = length(intersect(de.genes$`SCZ_DLPFC_CMC-Control_DLPFC_CMC.MATCHED.UP`, 
                        de.genes$`SCZ_DLPFC_HBCC-Control_DLPFC_HBCC.MATCHED.UP`))
  l3 = length(unique(Reduce(c, de.genes)))
  matched = (l1+l2)/l3
  n.matched = l1+l2

  rep.rate = c(rep.rate, list(data.frame(fdr = 10^pval, 
                                         wo.qsva = wo.qsva, w.qsva = w.qsva, matched = matched,
                                         n.wo.qsva = n.wo.qsva, n.w.qsva = n.w.qsva, n.matched = n.matched)))
}
rep.rate = rbindlist(rep.rate)
rep.rate %>%
  dplyr::select(fdr, wo.qsva, w.qsva, matched) %>%
  tidyr::gather(method, rep.rate, -fdr) %>%
  dplyr::mutate(rep.rate = rep.rate * 100) %>%
  dplyr::left_join(rep.rate %>%
                     dplyr::select(fdr, n.wo.qsva, n.w.qsva, n.matched) %>%
                     tidyr::gather(method, nOverlap, -fdr) %>%
                     dplyr::mutate(method = gsub('n.','',method))) %>%
  dplyr::filter(fdr <= 1e-1) %>%
  ggplot(aes(x=fdr, y=rep.rate, color = method)) + geom_point(aes(size = nOverlap)) + geom_line() + scale_x_log10() + xlab('FDR') + ylab('Replication Rate (%)') + ggtitle('CMC - HBCC - DLPFC')
```
Replication rate based on qvalue
```{r anlz.results1}
rep.rate = list()
for (pval in c(-1, -1.30103, -2)){
  de.genes = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC') %>%
    plyr::dlply(.(Model), .fun = function(x){
      x$ensembl_gene_id[x$adj.P.Val <= 10^pval] %>% unique()
    })

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC', Model == 'CMC', ensembl_gene_id %in% de.genes$HBCC) %>%
    dplyr::select(P.Value) %>%
    qvalue::qvalue(fdr.level = 10^pval)
  rs = data.frame(method = 'wo.qsva', ref.data = 'HBCC', rep.data = 'CMC',
                  rep.rate = 1-tmp$pi0, 
                  ncount = sum(tmp$significant))

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC', Model == 'HBCC', ensembl_gene_id %in% de.genes$CMC) %>%
    dplyr::select(P.Value) %>%
    qvalue::qvalue(fdr.level = 10^pval)
  rs = rbind(rs,
             data.frame(method = 'wo.qsva', ref.data = 'CMC', rep.data = 'HBCC',
                        rep.rate = 1-tmp$pi0, 
                        ncount = sum(tmp$significant)))

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC', Model == 'CMC_qSVA', ensembl_gene_id %in% de.genes$HBCC_qSVA) %>%
    dplyr::select(P.Value) %>%
    qvalue::qvalue(fdr.level = 10^pval)
  rs = rbind(rs,
             data.frame(method = 'w.qsva', ref.data = 'HBCC', rep.data = 'CMC',
                        rep.rate = 1-tmp$pi0, 
                        ncount = sum(tmp$significant)))

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC-Control_DLPFC', Model == 'HBCC_qSVA', ensembl_gene_id %in% de.genes$CMC_qSVA) %>%
    dplyr::select(P.Value) %>%
    qvalue::qvalue(fdr.level = 10^pval)
  rs = rbind(rs,
             data.frame(method = 'w.qsva', ref.data = 'CMC', rep.data = 'HBCC',
                        rep.rate = 1-tmp$pi0, 
                        ncount = sum(tmp$significant)))

  de.genes = diff.exp %>%
    dplyr::filter(Comparison %in% c('SCZ_DLPFC_CMC-Control_DLPFC_CMC', 'SCZ_DLPFC_HBCC-Control_DLPFC_HBCC')) %>%
    plyr::dlply(.(Comparison), .fun = function(x){
      x$ensembl_gene_id[x$adj.P.Val <= 10^pval] %>% unique()
    })

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC_CMC-Control_DLPFC_CMC', 
                  ensembl_gene_id %in% de.genes$`SCZ_DLPFC_HBCC-Control_DLPFC_HBCC`) %>%
    dplyr::select(P.Value)

  if (dim(tmp)[1] != 0){
    tmp %>%
      qvalue::qvalue(fdr.level = 10^pval)
    rs = rbind(rs,
               data.frame(method = 'matched', ref.data = 'HBCC', rep.data = 'CMC',
                          rep.rate = 1-tmp$pi0, 
                          ncount = sum(tmp$significant)))
  } else {
    rs = rbind(rs,
               data.frame(method = 'matched', ref.data = 'HBCC', rep.data = 'CMC',
                          rep.rate = 0, 
                          ncount = 0))
  }

  tmp = diff.exp %>%
    dplyr::filter(Comparison == 'SCZ_DLPFC_HBCC-Control_DLPFC_HBCC', 
                  ensembl_gene_id %in% de.genes$`SCZ_DLPFC_CMC-Control_DLPFC_CMC`) %>%
    dplyr::select(P.Value) %>%
    qvalue::qvalue(fdr.level = 10^pval)
  rs = rbind(rs,
             data.frame(method = 'matched', ref.data = 'CMC', rep.data = 'HBCC',
                        rep.rate = 1-tmp$pi0, 
                        ncount = sum(tmp$significant)))

  rs$fdr = 10^pval
  rep.rate = c(rep.rate, list(rs))
}
rep.rate = rbindlist(rep.rate)
rep.rate %>%
  ggplot(aes(x=fdr, y=rep.rate*100, color = method)) + geom_point(aes(size = ncount)) + geom_line() + scale_x_log10() + xlab('FDR') + ylab('Replication Rate (%)') + facet_grid(rep.data~., scales = 'free_y')
```