### 2017-10-09 code/Rmd/networkAssociationTest.Rmd
Added markdown

---
title: "Function to find association b/w n significant eigen genes with Diagnosis, Diagnosis.Gender, ApoE"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./networkAssociationTest.Rmd",
                                 parentId = "syn10306449",
                                 entityName = 'Network Association Tests')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

# Load Libraries
library(CovariateAnalysis) #devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(synapseClient)
library(knitr)
library(githubr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

```{r synapse.parameters, include=FALSE}
parentId = 'syn10306449';
activityName = 'Network associaiton statistics';

thisFileName <- 'networkAssociationTest.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_networks')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Data download
```{r download.data, include=FALSE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10338156')@values
all.used.ids = c('syn10338156')

MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$GeneID))

COVARIATES = downloadFile('syn10160582')
all.used.ids = c(all.used.ids, 'syn10160582')
COVARIATES$BrainRegion[COVARIATES$BrainRegion == 'CER'] = 'CBE'

FactorCovariates = c("individualIdentifier", "Study", "BrainRegion.Diagnosis", "BrainRegion", 
                     "Diagnosis", "Gender", "Batch", "BrainRegion.ApoE", "BrainRegion.Diagnosis.Gender")
ContCovariates = c("AOD", "PMI", "RIN", "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES", "PCT_INTERGENIC_BASES",
                   "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES", "RIN2") 
COVARIATES[,FactorCovariates] = lapply(COVARIATES[, FactorCovariates], as.factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[, ContCovariates], as.numeric)

EXPR = downloadFile('syn10160584')
all.used.ids = c(all.used.ids, 'syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL

MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values
all.used.ids = c(all.used.ids, 'syn10339153')
MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values
all.used.ids = c(all.used.ids, 'syn10492048')

MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
all.used.ids = c(all.used.ids, 'syn10380432')
```

```{r fxn}
# Function to find association statistics
findAssociationStatistics <- function(x, covars, exprs, frmla){

  covar = dplyr::filter(covars, BrainRegion == unique(x$brainRegion))
  expr = exprs[x$GeneID,covar$SampleID]

  result = list( assoc.stat = data.frame(total.eff.sz = 0,
                                         total.pval = 1,
                                         ngenes = length(unique(x$GeneID)),
                                         npc = 0,
                                         npc.sig = 0,
                                         pc1.eff.sz = 0,
                                         pc1.fdr = 1),
                 anova.stat = data.frame() )

  if (dim(expr)[1] <= 1){
    return(result)
  }

  # Function to get anova association stats
  getAnovaAssociationStats <- function(dat, frmla){
    mdl = as.data.frame(anova(lm(formula = frmla, data = dat)))
    mdl$fdr = p.adjust(mdl[,'Pr(>F)'])
    eff.sz = mdl['Diagnosis','Sum Sq']/sum(mdl[,'Sum Sq'], na.rm = T)
    fdr = mdl['Diagnosis','fdr']
    return(data.frame(eff.sz = eff.sz, fdr = fdr))
  }

  # Function to get PCs
  getPCs <- function(expr){
    expr = scale(t(scale(t(expr))))
    pc = prcomp(expr, center = F, scale. = F)
    pc.sdev = pc$sdev^2/sum(pc$sdev^2)
    ind = which(pc.sdev >= 0.01)
    pc.sdev = pc.sdev[ind]
    pc.rot = pc$rotation[,ind, drop = F]

    return(list(sdev = pc.sdev, rot = pc.rot))
  }

  # Function to get PCs
  bootstrapAssociationStats <- function(expr, covar, frmla){
    pc = getPCs(expr)

    anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
      dat = cbind(data.frame(PC = x), covar)
      getAnovaAssociationStats(dat, frmla)
    }, covar, frmla) %>%
      rbindlist(use.names = T, fill = T)
    anova.stat$fve = pc$sdev

    total.eff.sz = dplyr::filter(anova.stat, fdr <= 0.05) %>%
      dplyr::mutate(ws = eff.sz * fve) %>%
      dplyr::select(ws)

    sum(total.eff.sz$ws, na.rm = T)
  }

  pc = getPCs(expr)

  anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
    dat = cbind(data.frame(PC = x), covar)
    getAnovaAssociationStats(dat, frmla)
  }, covar, frmla) %>%
    rbindlist(use.names = T, fill = T)
  anova.stat$fve = pc$sdev

  total.eff.sz = dplyr::filter(anova.stat, fdr <= 0.05) %>%
    dplyr::mutate(ws = eff.sz * fve) %>%
    dplyr::select(ws) 
  total.eff.sz = sum(total.eff.sz$ws, na.rm = T)

  rn.eff.sz = plyr::laply(1:100, .fun = function(i, expr, l, covar, frmla){
    tryCatch({ 
      bootstrapAssociationStats(expr[sample(1:dim(expr)[1], l), ], covar, frmla) 
    }, error = function(e){ 0 })
  }, exprs[,covar$SampleID], length(x$GeneID), covar, frmla, .parallel = T)

  result = data.frame(total.eff.sz = total.eff.sz,
                      total.pval = sum(rn.eff.sz >= total.eff.sz, na.rm = T)/length(rn.eff.sz),
                      ngenes = length(unique(x$GeneID)),
                      npc = dim(anova.stat)[1],
                      npc.sig = sum(anova.stat$fdr <= 0.05),
                      pc1.eff.sz = anova.stat$eff.sz[1],
                      pc1.fdr = anova.stat$fdr[1])

  return(list(assoc.stat = result, anova.stat = anova.stat))
}

# Function to get module overlap plots
getOverlapPlots <- function(imp.mods, MOD.MEMBERS, MODULE.OVRLP){
    mdl.ovrlp = matrix(0, length(unique(imp.mods$ModuleNameFull)),length(unique(imp.mods$ModuleNameFull)))
    rownames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    colnames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    for( i in rownames(mdl.ovrlp)){
      for( j in colnames(mdl.ovrlp)){
        mdl.ovrlp[i,j] = length(intersect(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]]))/length(unique(c(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]])))
      }
    }

    mdl.ovrlp.pval = MODULE.OVRLP %>%
      dplyr::filter(ModuleNameFull %in% unique(imp.mods$ModuleNameFull), category %in% unique(imp.mods$ModuleNameFull)) %>%
      dplyr::select(ModuleNameFull, category, fisherPval) %>%
      igraph::graph_from_data_frame(directed=F) %>%
      igraph::as_adj(attr = 'fisherPval')
    mdl.ovrlp.pval = mdl.ovrlp.pval/2
    mdl.ovrlp.log = matrix('*', dim(mdl.ovrlp.pval)[1], dim(mdl.ovrlp.pval)[2])
    mdl.ovrlp.log[as.matrix(mdl.ovrlp.pval) > 0.05] = ''
    rownames(mdl.ovrlp.log) = rownames(mdl.ovrlp.pval)
    colnames(mdl.ovrlp.log) = colnames(mdl.ovrlp.pval)
    mdl.ovrlp.log = mdl.ovrlp.log[rownames(mdl.ovrlp), colnames(mdl.ovrlp)]

    Heatmap(mdl.ovrlp, col = colorRamp2(seq(0,1,length.out = 5), viridis::viridis(5)),
            name = 'JacDist', cell_fun = function(j,i,x,y,wifht,height,fill){
              grid.text(mdl.ovrlp.log[i,j] %>% format(digits = 1), x, y)
            })
  }
```

### Associaiton with Diagnosis
```{r Dx, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(Dx, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()
```
Top 5 modules per method
```{r Dx.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(Dx.gender, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()
```
Top 5 modules per method
```{r Dx.gender.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  if(dim(imp.mods)[1] != 0 ){
    getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
  }
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.gender.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  if(dim(imp.mods)[1] != 0 ){
    print(summary(factor(imp.mods$method)))
    getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
  }
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
# draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

### Associaiton with ApoE
```{r ApoE, fig.height=5, fig.width=10}
# Find assocition b/w ApoE category with n eigen genes
FORMULA = as.formula('PC~BrainRegion.ApoE+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.apoe = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

ApoE = lapply(assoc.stat.apoe, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(ApoE, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```

### Select information rich module (based on Dx association, MAGMA enrichment, dbgap enrichment and AD related pathway enrichment)
```{r imp.mdl, fig.height=7, fig.width=12}
ad.magma = MAGMA.ENRICH %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(P),
                    rnk = rank(rank(-log10(fdr))+rank(BETA)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(SET) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

ad.dbgap = MODULE.ENRICH %>%
  dplyr::filter(geneSet == 'dbgap', category == 'Alzheimer.Disease') %>%
  unique() %>%
  dplyr::left_join(Dx %>% dplyr::select(ModuleName, brainRegion, ModuleNameFull, method)) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(fisherPval),
                    rnk = rank(rank(-log10(fdr))+rank(fisherOR)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk)) %>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

other.ad.pathways = c("alzheimer_disease", "Alzheimer.s.disease_Homo.sapiens_hsa05010", 
                      "Alzheimers.Disease_Homo.sapiens_WP2059",
                      "Alzheimer.disease.amyloid.secretase.pathway_Homo.sapiens_P00003",
                      "Alzheimer.s_disease", "Alzheimer.Disease")
ad.others = MODULE.ENRICH %>%
  dplyr::filter(geneSet != 'dbgap', category %in% other.ad.pathways) %>%
  unique() %>%
  dplyr::left_join(Dx %>% dplyr::select(ModuleName, brainRegion, ModuleNameFull, method)) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(fisherPval),
                    rnk = rank(rank(-log10(fdr))+rank(fisherOR)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

# Combine and rank all the information
scr1 = Dx %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(rnk = rank(rank(total.eff.sz) + rank(-log10(total.fdr))),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(pc1.fdr))
  }) %>%
  dplyr::mutate(Dx = rank(rnk),
                Dx = Dx/max(Dx)) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, ModuleName, Dx)

scr2 = ad.magma %>%
  dplyr::mutate(magma = rank(rnk),
                magma = magma/max(magma)) %>%
  dplyr::rename(ModuleNameFull = SET) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, magma)

scr3 = ad.dbgap %>%
  dplyr::mutate(dbgap = rank(rnk),
                dbgap = dbgap/max(dbgap)) %>%
  dplyr::select(ModuleNameFull, ModuleName, brainRegion, method, dbgap)

scr4 = ad.others %>%
  dplyr::mutate(other = rank(rnk),
                other = other/max(other)) %>%
  dplyr::select(ModuleNameFull, ModuleName, brainRegion, method, other)

all.rank = plyr::join_all(list(scr1, scr2, scr3, scr4), type = 'full')
all.rank[is.na(all.rank)] = 0
all.rank = all.rank %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x = x %>%
      dplyr::mutate(rnk = rank(Dx+magma+dbgap+other),
                    rnk = rnk/max(rnk))
    print(paste('BrainRegion:', unique(x$brainRegion)))
    print(paste('Correlation with Dx', round(cor(x$rnk, x$Dx),2)))
    print(paste('Correlation with magma', round(cor(x$rnk, x$magma),2)))
    print(paste('Correlation with dbgap', round(cor(x$rnk, x$dbgap),2)))
    print(paste('Correlation with others', round(cor(x$rnk, x$other),2)))
    return(x)
  })

ggplot(all.rank, aes(x = Dx, y = magma))+geom_point()+facet_grid(.~brainRegion)

ggplot(all.rank, aes(x = dbgap, y = other))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')
ggplot(all.rank, aes(x = dbgap, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')

ggplot(all.rank, aes(x = other, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')

ggplot(tmp, aes(x = method, y = rnk)) + geom_boxplot() + facet_grid(.~brainRegion) + theme(axis.text.x = element_text(angle=45, hjust = 1)) + coord_flip() + geom_jitter()

```
Top 5 modules per method
```{r imp.mdl.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, rnk) %>%
    dplyr::top_n(5, Dx) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r imp.mdl.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, rnk) %>%
    dplyr::top_n(25, Dx) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

### Modules missing enrichment scores
```{r missing, include=FALSE}
writeLines('Modules missing enrich enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MODULE.ENRICH$ModuleNameFull))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()

writeLines('Modules missing MAGMA enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MAGMA.ENRICH$Set))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()
```

### Store results in synapse
```{r syn.store, include=FALSE}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn10831307')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Dx.gender, file = 'AssociaitonStatWithDxGender.tsv', sep = '\t', quote = F, row.names = F)
OBJ1 = File('AssociaitonStatWithDxGender.tsv', name = 'Association with Dx.Gender', parentId = 'syn10831307')
OBJ1 = synStore(OBJ1, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(ApoE, file = 'AssociaitonStatWithApoE4.tsv', sep = '\t', quote = F, row.names = F)
OBJ2 = File('AssociaitonStatWithApoE4.tsv', name = 'Association with ApoE4', parentId = 'syn10831307')
OBJ2 = synStore(OBJ2, used = all.used.ids, executed = thisFile, activityName = activityName)

save(list = c('assoc.stat.diagnosis', 'assoc.stat.diagnosis.gender', 'assoc.stat.apoe'), file = 'AllAssociationResults.RData')
OBJ3 = File('AllAssociationResults.RData', name = 'All Assocition Statistics (RData)', parentId = 'syn10831307')
OBJ3 = synStore(OBJ3, used = all.used.ids, executed = thisFile, activityName = activityName)
```