### 2017-10-10 code/Rmd/networkAssociationTest_metaModules.Rmd
Temp commit

---
title: "Function to find association b/w n significant eigen genes of meta modules with Diagnosis, Diagnosis.Gender, ApoE"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./networkAssociationTest_metaModules.Rmd",
                                 parentId = "syn10306449",
                                 entityName = 'Network Association Tests (aggregated modules)')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

# Load Libraries
library(CovariateAnalysis) #devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(synapseClient)
library(knitr)
library(githubr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE}
parentId = 'syn10306449';
activityName = 'Network associaiton statistics';

thisFileName <- 'networkAssociationTest_metaModules.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_networks')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Data download
```{r download.data, include=FALSE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10915669')@values
all.used.ids = c('syn10915669')

MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$GeneID))

COVARIATES = downloadFile('syn10160582')
all.used.ids = c(all.used.ids, 'syn10160582')
COVARIATES$BrainRegion[COVARIATES$BrainRegion == 'CER'] = 'CBE'

FactorCovariates = c("individualIdentifier", "Study", "BrainRegion.Diagnosis", "BrainRegion", 
                     "Diagnosis", "Gender", "Batch", "BrainRegion.ApoE", "BrainRegion.Diagnosis.Gender")
ContCovariates = c("AOD", "PMI", "RIN", "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES", "PCT_INTERGENIC_BASES",
                   "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES", "RIN2") 
COVARIATES[,FactorCovariates] = lapply(COVARIATES[, FactorCovariates], as.factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[, ContCovariates], as.numeric)

EXPR = downloadFile('syn10160584')
all.used.ids = c(all.used.ids, 'syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL
```

```{r fxn}
# Function to find association statistics
findAssociationStatistics <- function(x, covars, exprs, frmla){

  covar = dplyr::filter(covars, BrainRegion == unique(x$brainRegion))
  expr = exprs[x$GeneID,covar$SampleID]

  result = list( assoc.stat = data.frame(total.eff.sz = 0,
                                         total.pval = 1,
                                         ngenes = length(unique(x$GeneID)),
                                         npc = 0,
                                         npc.sig = 0,
                                         pc1.eff.sz = 0,
                                         pc1.fdr = 1),
                 anova.stat = data.frame() )

  if (dim(expr)[1] <= 1){
    return(result)
  }

  # Function to get anova association stats
  getAnovaAssociationStats <- function(dat, frmla){
    mdl = as.data.frame(anova(lm(formula = frmla, data = dat)))
    mdl$fdr = p.adjust(mdl[,'Pr(>F)'])
    eff.sz = mdl['Diagnosis','Sum Sq']/sum(mdl[,'Sum Sq'], na.rm = T)
    fdr = mdl['Diagnosis','fdr']
    return(data.frame(eff.sz = eff.sz, fdr = fdr))
  }

  # Function to get PCs
  getPCs <- function(expr){
    expr = scale(t(scale(t(expr))))
    pc = prcomp(expr, center = F, scale. = F)
    pc.sdev = pc$sdev^2/sum(pc$sdev^2)
    ind = which(pc.sdev >= 0.01)
    pc.sdev = pc.sdev[ind]
    pc.rot = pc$rotation[,ind, drop = F]

    return(list(sdev = pc.sdev, rot = pc.rot))
  }

  # Function to get PCs
  bootstrapAssociationStats <- function(expr, covar, frmla){
    pc = getPCs(expr)

    anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
      dat = cbind(data.frame(PC = x), covar)
      getAnovaAssociationStats(dat, frmla)
    }, covar, frmla) %>%
      rbindlist(use.names = T, fill = T)
    anova.stat$fve = pc$sdev

    total.eff.sz = dplyr::filter(anova.stat, fdr <= 0.05) %>%
      dplyr::mutate(ws = eff.sz * fve) %>%
      dplyr::select(ws)

    sum(total.eff.sz$ws, na.rm = T)
  }

  pc = getPCs(expr)

  anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
    dat = cbind(data.frame(PC = x), covar)
    getAnovaAssociationStats(dat, frmla)
  }, covar, frmla) %>%
    rbindlist(use.names = T, fill = T)
  anova.stat$fve = pc$sdev

  total.eff.sz = dplyr::filter(anova.stat, fdr <= 0.05) %>%
    dplyr::mutate(ws = eff.sz * fve) %>%
    dplyr::select(ws) 
  total.eff.sz = sum(total.eff.sz$ws, na.rm = T)

  rn.eff.sz = plyr::laply(1:100, .fun = function(i, expr, l, covar, frmla){
    tryCatch({ 
      bootstrapAssociationStats(expr[sample(1:dim(expr)[1], l), ], covar, frmla) 
    }, error = function(e){ 0 })
  }, exprs[,covar$SampleID], length(x$GeneID), covar, frmla, .parallel = T)

  result = data.frame(total.eff.sz = total.eff.sz,
                      total.pval = sum(rn.eff.sz >= total.eff.sz, na.rm = T)/length(rn.eff.sz),
                      ngenes = length(unique(x$GeneID)),
                      npc = dim(anova.stat)[1],
                      npc.sig = sum(anova.stat$fdr <= 0.05),
                      pc1.eff.sz = anova.stat$eff.sz[1],
                      pc1.fdr = anova.stat$fdr[1])

  return(list(assoc.stat = result, anova.stat = anova.stat))
}

# Function to get module overlap plots
getOverlapPlots <- function(imp.mods, MOD.MEMBERS){
    mdl.ovrlp = matrix(0, length(unique(imp.mods$ModuleNameFull)),length(unique(imp.mods$ModuleNameFull)))
    rownames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    colnames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    for( i in rownames(mdl.ovrlp)){
      for( j in colnames(mdl.ovrlp)){
        mdl.ovrlp[i,j] = length(intersect(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]]))/length(unique(c(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]])))
        fisherEnrichment()
      }
    }

    mdl.ovrlp.pval = MODULE.OVRLP %>%
      dplyr::filter(ModuleNameFull %in% unique(imp.mods$ModuleNameFull), category %in% unique(imp.mods$ModuleNameFull)) %>%
      dplyr::select(ModuleNameFull, category, fisherPval) %>%
      igraph::graph_from_data_frame(directed=F) %>%
      igraph::as_adj(attr = 'fisherPval')
    mdl.ovrlp.pval = mdl.ovrlp.pval/2
    mdl.ovrlp.log = matrix('*', dim(mdl.ovrlp.pval)[1], dim(mdl.ovrlp.pval)[2])
    mdl.ovrlp.log[as.matrix(mdl.ovrlp.pval) > 0.05] = ''
    rownames(mdl.ovrlp.log) = rownames(mdl.ovrlp.pval)
    colnames(mdl.ovrlp.log) = colnames(mdl.ovrlp.pval)
    mdl.ovrlp.log = mdl.ovrlp.log[rownames(mdl.ovrlp), colnames(mdl.ovrlp)]

    Heatmap(mdl.ovrlp, col = colorRamp2(seq(0,1,length.out = 5), viridis::viridis(5)),
            name = 'JacDist', cell_fun = function(j,i,x,y,wifht,height,fill){
              grid.text(mdl.ovrlp.log[i,j] %>% format(digits = 1), x, y)
            })
  }
```
### Associaiton with Diagnosis
```{r Dx, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  }) %>%
  dplyr::mutate(total.sig = total.fdr <= 0.05)

p = ggplot(Dx, aes(x = ngenes, y = total.eff.sz, color = brainRegion)) + geom_point()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = as.factor(npc.sig), y = total.eff.sz, color = brainRegion)) + geom_boxplot()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = brainRegion, y = total.eff.sz)) + geom_boxplot() + geom_jitter(aes(color = total.sig))
p = p + coord_flip() + xlab('Brain Region') + ylab('Total Effect Size')
# p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = brainRegion, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + xlab('Brain Region') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05),
                   ntotal = sum(total.fdr <= 1)) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```
### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=10, fig.width=10}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  }) %>%
  dplyr::mutate(total.sig = total.fdr <= 0.05)

p = ggplot(Dx.gender, aes(x = ngenes, y = total.eff.sz, color = brainRegion)) + geom_point()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = as.factor(npc.sig), y = total.eff.sz, color = brainRegion)) + geom_boxplot()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = brainRegion, y = total.eff.sz)) + geom_boxplot() + geom_jitter(aes(color = total.sig))
p = p + coord_flip() + xlab('Brain Region') + ylab('Total Effect Size')
# p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = brainRegion, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + xlab('Brain Region') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05),
                   ntotal = sum(total.fdr <= 1)) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Comparison b/w Dx and Dx.gender')
tmp = dplyr::left_join(Dx %>%
                         dplyr::select(ModuleName, brainRegion, total.eff.sz, total.fdr) %>%
                         dplyr::rename(Dx = total.eff.sz, Dx.fdr = total.fdr),
                       Dx.gender %>%
                         dplyr::select(ModuleName, brainRegion, total.eff.sz, total.fdr) %>%
                         dplyr::rename(Dx.gender = total.eff.sz, Dx.gender.fdr = total.fdr))
p = ggplot(tmp, aes(x = Dx, y = Dx.gender, color = -log10(Dx.fdr+1e-100))) + geom_point() + facet_grid(.~brainRegion) + geom_smooth(method = 'lm')
p = p +scale_color_continuous(low = 'white', high = 'red')
p
```
### Enrichment for differential expression
```{r Dx.enrich, fig.height=10, fig.width=10}
# Get differential expression gene list from synapse
load(synGet('syn10496554')@filePath)
all.used.ids = c(all.used.ids, 'syn10496554')
names(amp.ad.de.geneSets) = gsub('CER','CBE', names(amp.ad.de.geneSets))

# Perform enrichment analysis
MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$external_gene_name))
Dx.enrich = plyr::ldply(unique(MODULES$brainRegion), .fun = function(x){
  modMembers = MOD.MEMBERS[grep(x, names(MOD.MEMBERS))]
  deGenes = amp.ad.de.geneSets[grep(x, names(amp.ad.de.geneSets))]
  backgroundGenes = unique(MODULES$external_gene_name)

  enrichResults = plyr::ldply(modMembers, .fun = function(y, deGenes, backgroundGenes){
    plyr::ldply(deGenes, .fun = function(z, y, backgroundGenes){
      fisherEnrichment(y,z,backgroundGenes)
    }, y, backgroundGenes, .id = 'Dx.GeneSet')
  }, deGenes, backgroundGenes, .id = 'ModuleNameFull') %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, .id = 'brainRegion')

writeLines('Comparison b/w Dx and Dx.gender')
Dx.enrich1 = Dx.enrich %>%
  dplyr::select(-Genes) %>%
  tidyr::separate(Dx.GeneSet, c('brainRegion', 'Comparison', 'Gender', 'Direction'), sep = '\\.')
Dx.enrich1$Direction[is.na(Dx.enrich1$Direction)] = Dx.enrich1$Gender[is.na(Dx.enrich1$Direction)]
Dx.enrich1$Gender = gsub('UP', 'ALL', Dx.enrich1$Gender)
Dx.enrich1$Gender = gsub('DOWN', 'ALL', Dx.enrich1$Gender)
Dx.enrich1 = dplyr::filter(Dx.enrich1, Comparison == 'AD-CONTROL', fdr <= 0.05)
p = ggplot(Dx.enrich1, aes(x = Gender, y = -log10(fdr))) + geom_boxplot() 
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free') + geom_jitter(aes(color = Direction))
pl = list();pl[[1]] = p

p = ggplot(Dx.enrich1, aes(x = Gender, y = Odds.Ratio)) + geom_boxplot() 
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free') + geom_jitter(aes(color = Direction))
pl[[2]] = p
multiplot(plotlist = pl, cols = 1)
```
### Enrichment for gene sets
```{r pathway.enrich, results='asis'}
# Download all related genesets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c(all.used.ids, 'syn4867851')

# Download cell type specific genesets
load(synGet('syn9818080')@filePath)
all.used.ids = c(all.used.ids, 'syn9818080')

backgroundGenes = unique(MODULES$external_gene_name)
geneSet.ref = c(GeneSets[c('KEGG_2016', 'GO_Biological_Process_2015', 'Reactome_2016', 'WikiPathways_2016')],
                list(CellType = CellTypeList) )%>%
  lapply(function(x){
    lapply(x, function(y){
      sapply(y, function(y) str_split(y, '\\,')[[1]][1])
    })
  }) %>%
  CovariateAnalysis::filterGeneSets(backgroundGenes) %>%
  do.call(c,.)

# Perform enrichment analysis
Pathway.enrich = plyr::ddply(MODULES, .(ModuleNameFull), .fun = function(mod, geneSetRef, backgroundGenes){
  modMembers = unique(mod$external_gene_name)

  enrichResults = plyr::ldply(geneSetRef, .fun = function(y, modMembers, backgroundGenes){
    fisherEnrichment(y, modMembers, backgroundGenes)
  }, modMembers, backgroundGenes, .id = 'GeneSetName', .parallel = T, 
  .paropts = list(.packages = c('CovariateAnalysis'))) %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, 
.id = 'ModuleNameFull', geneSet.ref, backgroundGenes)

# Get AD specific pathways
other.ad.pathways = c("KEGG_2016.Alzheimer's disease_Homo sapiens_hsa05010",
                      "WikiPathways_2016.Alzheimers Disease_Homo sapiens_WP2059")

writeLines('AD related pathways enriched an fdr of 0.05')
ad.enrich = Pathway.enrich %>%
  dplyr::select(ModuleNameFull, GeneSetName, Odds.Ratio, fdr) %>%
  dplyr::filter(GeneSetName %in% other.ad.pathways, fdr <= 0.05) %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(ModuleNameFull, brainRegion)) %>%
  unique()
kable(ad.enrich)
```
### Find overlap between modules in different brainregion
```{r module.overlap, fig.height=5, fig.width=10}
# FisherEnrichment b/w modules
MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$external_gene_name))
MOD.OVRLP = plyr::ldply(MOD.MEMBERS, .fun = function(x, modMembers, backgroundGenes){
  plyr::ldply(modMembers, .fun = function(y, x, backgroundGenes){
    fisherEnrichment(y, x, backgroundGenes)
  }, x, backgroundGenes, .parallel = T, .paropts = list(.packages = c('CovariateAnalysis')), .id = 'MOD1') %>%
    dplyr::rename(MOD1 = ModuleNameFull)
}, MOD.MEMBERS, backgroundGenes, .id = 'MOD2') %>%
  dplyr::select(-Genes) %>%
  dplyr::rename(MOD2 = ModuleNameFull)
```
### Interesting modules
```{r int.modules, fig.height=7, fig.width=7, results='asis'}
# Wrt case control
tmp = inner_join(Dx %>% 
                   dplyr::filter(total.fdr <= 0.05) %>%
                   dplyr::select(ModuleNameFull, brainRegion, total.eff.sz, total.fdr) %>%
                   dplyr::rename(assoc.eff.sz = total.eff.sz, assoc.fdr = total.fdr),
                 Dx.enrich1 %>%
                   dplyr::filter(fdr <= 0.05, Gender == 'ALL') %>%
                   dplyr::select(ModuleNameFull, brainRegion, Direction, Odds.Ratio, fdr) %>%
                   dplyr::rename(diffexp.fdr = fdr, diffexp.or = Odds.Ratio)) %>%
  left_join(ad.enrich %>%
              dplyr::select(ModuleNameFull, fdr, brainRegion) %>%
              dplyr::rename(ad.pathway.fdr = fdr))
writeLines('Imporatant Modules for case control comparison')
kable(tmp)

writeLines('Up regulated modules')
tmp1 = filter(tmp, Direction == 'UP')
tmp1 = filter(MOD.OVRLP, MOD1 %in% tmp1$ModuleNameFull, MOD2 %in% tmp1$ModuleNameFull)
g = igraph::graph_from_data_frame(tmp1)
tmp1.or = igraph::as_adj(g, attr = 'Odds.Ratio') %>% data.matrix
tmp1.or[is.infinite(tmp1.or)] = 0

tmp1.fdr = igraph::as_adj(g, attr = 'pval') %>% data.matrix()
tmp1.fdr = tmp1.fdr <= 0.05
tmp1.fdr[tmp1.fdr] = '*'
tmp1.fdr[tmp1.fdr == 'FALSE'] = ''

h = ComplexHeatmap::Heatmap(log2(tmp1.or+1e-2), col = viridis::viridis(5), 
                        cell_fun = function(j,i,x,y,wifht,height,fill){
                          grid.text(tmp1.fdr[i,j] %>% format(digits = 1), x, y)
                        }, name = 'log2(OR)')
draw(h, heatmap_legend_side = 'left')

dplyr::filter(Pathway.enrich, 
              GeneSetName %in% grep('CellType', unique(Pathway.enrich$GeneSetName), value = T),
              fdr <= 0.05, 
              ModuleNameFull %in% rownames(tmp1.or)) %>%
  dplyr::select(-Genes)

writeLines('Down regulated modules')
tmp1 = filter(tmp, Direction == 'DOWN')
tmp1 = filter(MOD.OVRLP, MOD1 %in% tmp1$ModuleNameFull, MOD2 %in% tmp1$ModuleNameFull)
g = igraph::graph_from_data_frame(tmp1)
tmp1.or = igraph::as_adj(g, attr = 'Odds.Ratio') %>% data.matrix
tmp1.or[is.infinite(tmp1.or)] = 0

tmp1.fdr = igraph::as_adj(g, attr = 'pval') %>% data.matrix()
tmp1.fdr = tmp1.fdr <= 0.05
tmp1.fdr[tmp1.fdr] = '*'
tmp1.fdr[tmp1.fdr == 'FALSE'] = ''

h = ComplexHeatmap::Heatmap(log2(tmp1.or+1e-2), col = viridis::viridis(5), 
                        cell_fun = function(j,i,x,y,wifht,height,fill){
                          grid.text(tmp1.fdr[i,j] %>% format(digits = 1), x, y)
                        }, name = 'log2(OR)')
draw(h, heatmap_legend_side = 'left')

dplyr::filter(Pathway.enrich, 
              GeneSetName %in% grep('CellType', unique(Pathway.enrich$GeneSetName), value = T),
              fdr <= 0.05, 
              ModuleNameFull %in% rownames(tmp1.or)) %>%
  dplyr::select(-Genes)

```

### Store results in synapse
```{r syn.store, include=FALSE, eval = F, cache=FALSE}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn11140497')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Dx.gender, file = 'AssociaitonStatWithDxGender.tsv', sep = '\t', quote = F, row.names = F)
OBJ1 = File('AssociaitonStatWithDxGender.tsv', name = 'Association with Dx.Gender', parentId = 'syn11140497')
OBJ1 = synStore(OBJ1, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Pathway.enrich, file = 'PathwayEnrichment.tsv', sep = '\t', quote = F, row.names = F)
OBJ2 = File('PathwayEnrichment.tsv', name = 'Pathway Enrichment', parentId = 'syn11140497')
OBJ2 = synStore(OBJ2, used = all.used.ids, executed = thisFile, activityName = activityName)

save(list = c('assoc.stat.diagnosis', 'assoc.stat.diagnosis.gender'), file = 'AllAssociationResults.RData')
OBJ3 = File('AllAssociationResults.RData', name = 'All Assocition Statistics (RData)', parentId = 'syn11140497')
OBJ3 = synStore(OBJ3, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Dx.enrich, file = 'DiagnosisEnrichment.tsv', sep = '\t', quote = F, row.names = F)
OBJ4 = File('DiagnosisEnrichment.tsv', name = 'Diagnosis Enrichment', parentId = 'syn11140497')
OBJ4 = synStore(OBJ4, used = all.used.ids, executed = thisFile, activityName = activityName)
```