### 2018-01-23 manuscriptFigures/diffExpFigure.Rmd
Added notebook to plot differential expression results

---
title: "Differential expression figures"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_notebook
---
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
# Load required libraries
library(CovariateAnalysis)
library(data.table)
library(plyr)
library(tidyverse)

library(synapseClient)
library(githubr)
library(knitr)

synapseLogin()

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Covariates
```{r technical.cov}
cov = downloadFile('syn10160582')
all.used.ids = c('syn10160582')
```
Total number of samples
```{r total.sample}
cov %>%
  dplyr::select(Study, SampleID, BrainRegion, Diagnosis) %>%
  dplyr::group_by(BrainRegion, Diagnosis) %>%
  dplyr::count() %>%
  tidyr::spread(Diagnosis, n) %>%
  kable()
```
Total number of samples per sex
```{r total.sample.sex}
cov %>%
  dplyr::select(Study, SampleID, BrainRegion, Diagnosis, Gender) %>%
  dplyr::group_by(BrainRegion, Gender) %>%
  dplyr::count() %>%
  tidyr::spread(Gender, n) %>%
  kable()
```
Different types of RNAs
```{r diff.rnas}
tmp = cov %>%
  dplyr::select(Study, PCT_CODING_BASES, PCT_INTERGENIC_BASES, 
                PCT_INTRONIC_BASES, PCT_RIBOSOMAL_BASES) %>%
  tidyr::gather(Feature, Value, -Study) %>%
  dplyr::group_by(Study, Feature) %>%
  dplyr::summarise(Value = mean(Value, na.rm = T)) %>%
  tidyr::spread(Feature, Value) %>%
  dplyr::group_by(Study) %>%
  dplyr::mutate(PCT_UTR_BASES = 1 - PCT_CODING_BASES - PCT_INTERGENIC_BASES - PCT_INTRONIC_BASES - PCT_RIBOSOMAL_BASES) %>%
  tidyr::gather(TypeOfRNA, Value, -Study)

p = ggplot(tmp, aes(x = Study, y= Value, fill = TypeOfRNA)) + geom_bar(stat = 'identity')
p = p + ylab('Fraction of aligned reads') + xlab('Study') + theme_bw()
p
```
### Differential Expression
```{r get.results}
dexp = downloadFile('syn11180450')
all.used.ids = c(all.used.ids, 'syn11180450')

ndexp = dplyr::filter(dexp, Model == 'Diagnosis') %>% 
  dplyr::group_by(Study, Tissue, Comparison, Gender) %>% 
  dplyr::summarise(UP = sum(adj.P.Val <= 0.05 & logFC >= 0), 
                   DOWN = sum(adj.P.Val <= 0.05 & logFC <= 0)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-Comparison, -Gender)
kable(ndexp)
```
pca plot
```{r pca}
expr = downloadFile('syn10160584') %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id','position'),sep = '\\.') %>%
  dplyr::filter(ensembl_gene_id %in% unique(dexp$ensembl_gene_id[dexp$adj.P.Val <= 0.05 & dexp$Model == 'Diagnosis']))

pc = prcomp(expr[,-(1:2)], scale. = T, center = T)

cov = data.frame(SampleID = colnames(expr)[-(1:2)],
           PC1 = pc$rotation[,1],
           PC2 = pc$rotation[,2]) %>%
  dplyr::left_join(cov)
p = ggplot(cov, aes(x = PC1, y = PC2, color = Study, shape = BrainRegion))
p = p + geom_point() + scale_shape_manual(values = 1:7)
p = p + theme(legend.position = 'top')
p
```

Volcano plots
```{r vp, fig.height=5, fig.width=10}
dexp$Direction = 'None'
dexp$Direction[dexp$logFC >= 0 & dexp$adj.P.Val <= 0.05] = 'Up'
dexp$Direction[dexp$logFC <= 0 & dexp$adj.P.Val <= 0.05] = 'Down'

p = ggplot(dexp %>%
             dplyr::filter(Model == 'Diagnosis'), 
           aes(x = logFC, y = -log10(adj.P.Val), color = Direction)) + geom_point(alpha = 0.1)
p = p + facet_grid(.~Study+Tissue, scales = 'free') + theme(legend.position = 'top') 
p = p + scale_color_manual(values = viridis::viridis(3))
# dev.off()
# svg(filename = 'VolcanoPlots.svg', height = 6, width = 12)
# print(p)
# dev.off()
p
```
Average expression
```{r aexp, fig.height=5, fig.width=10}
col = viridis::viridis(3)
p = ggplot(dexp%>%
             dplyr::filter(Model == 'Diagnosis'), 
           aes(x = AveExpr, y = logFC, color = Direction)) + geom_point(alpha = 0.1)
p = p + facet_grid(.~Study+Tissue, scales = 'free') + theme(legend.position = 'top') 
p = p + scale_color_manual(values = col)
p
```
Number of differentially expressed genes in each foldchange range (histogram)
```{r ndiff.range, fig.height=5, fig.width=10}
dexp$lfc1 = ifelse(dexp$logFC >= 0, dexp$logFC, NA)
dexp$lfc2 = ifelse(dexp$logFC <= 0, dexp$logFC, NA)
p = ggplot(dexp %>% dplyr::filter(adj.P.Val <= 0.05 & Model == 'Diagnosis'))
p = p + geom_histogram( aes(x = lfc1, y = ..count..), fill = "red")
p = p + geom_histogram( aes(x = lfc2, y = -..count..), fill = "green")
p = p + facet_grid(.~Study+Tissue, scales = 'free') + ylab('Gene count') + xlab('Log Fold Change')
p
```
### Overlap statistics
overlap between different brain regions
```{r ovlp.stats.up}
dexp.enrich.plots = dexp %>%
  dplyr::filter(Model %in% c('Diagnosis'), 
                adj.P.Val <= 0.05) %>%
  plyr::dlply(.(Direction), .fun = function(x){
    gr = x %>%
      plyr::dlply(.(Tissue), .fun = function(x){unique(x$ensembl_gene_id)}) %>%
      plyr::ldply(.,function(x, gs, backgroundGenes){
        plyr::ldply(gs, .fun = function(y,x,backgroundGenes){
          CovariateAnalysis::fisherEnrichment(x,y,backgroundGenes)
        }, x, backgroundGenes) %>%
          dplyr::rename(To = Tissue)
      }, ., background.genes) %>%
      dplyr::rename(From = Tissue) %>%
      dplyr::mutate(FDR = p.adjust(pval, method = 'BH')) %>%
      dplyr::select(From, To, Odds.Ratio, FDR, noverlap) %>%
      igraph::graph_from_data_frame()

    or.mat = igraph::as_adj(gr, type = 'both', attr = 'Odds.Ratio') %>% as.matrix()
    or.mat[is.infinite(or.mat)] = 0
    no.mat = igraph::as_adj(gr, type = 'both', attr = 'noverlap')

    if (unique(x$Direction) == 'Up'){
      col = c('white','red')
    } else {
      col = c('white','darkgreen')
    }
    ComplexHeatmap::Heatmap(as.matrix(or.mat), 
                            col= circlize::colorRamp2(c(0,40),col),
                            cell_fun = function(j, i, x, y, widht, height, fill){
                              no.mat[i,j] %>%
                                format(digit = 1) %>%
                                as.character() %>%
                                grid.text(x, y)
                            },
                            name = 'Odds Ratio',
                            column_title = paste(unique(x$Direction), 
                                                 'regulated'))
  })
purrr::walk(dexp.enrich.plots, function(x){ print(x) })
```

### Sex Specific Differential Expression
```{r ndiff.sex}
ndexp = dplyr::filter(dexp, Model == 'Diagnosis.Gender') %>% 
  dplyr::group_by(Study, Tissue, Comparison, Gender) %>% 
  dplyr::summarise(UP = sum(adj.P.Val <= 0.05 & logFC >= 0), 
                   DOWN = sum(adj.P.Val <= 0.05 & logFC <= 0)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-Comparison)
ndexp.female = dplyr::filter(ndexp, Gender == 'FEMALE') %>%
  dplyr::rename(FEMALE.UP = UP, FEMALE.DOWN = DOWN) %>%
  dplyr::select(-Gender)
ndexp.male = dplyr::filter(ndexp, Gender == 'MALE') %>%
  dplyr::rename(MALE.UP = UP, MALE.DOWN = DOWN) %>%
  dplyr::select(-Gender)
kable(dplyr::full_join(ndexp.female, ndexp.male, by = c('Tissue', 'Study')))
```
Volcano plots
```{r sex.vp, fig.height=10, fig.width=10}
p = ggplot(dexp %>%
             dplyr::filter(Model == 'Diagnosis.Gender'), 
           aes(x = logFC, y = -log10(adj.P.Val), color = Direction)) + geom_point(alpha = 0.1)
p = p + facet_grid(Gender+.~Study+Tissue, scales = 'free') + theme(legend.position = 'top') 
p = p + scale_color_manual(values = viridis::viridis(3))
# dev.off()
# svg(filename = 'VolcanoPlots.svg', height = 6, width = 12)
# print(p)
# dev.off()
p
```
Average expression
```{r sex.aexp, fig.height=10, fig.width=10}
col = viridis::viridis(3)
p = ggplot(dexp%>%
             dplyr::filter(Model == 'Diagnosis.Gender'), 
           aes(x = AveExpr, y = logFC, color = Direction)) + geom_point(alpha = 0.1)
p = p + facet_grid(Gender+.~Study+Tissue, scales = 'free') + theme(legend.position = 'top') 
p = p + scale_color_manual(values = col)
p
```
Number of differentially expressed genes in each foldchange range (histogram)
```{r sex.ndiff.range, fig.height=10, fig.width=10}
dexp$lfc1 = ifelse(dexp$logFC >= 0, dexp$logFC, NA)
dexp$lfc2 = ifelse(dexp$logFC <= 0, dexp$logFC, NA)
p = ggplot(dexp %>% dplyr::filter(adj.P.Val <= 0.05 & Model == 'Diagnosis.Gender'))
p = p + geom_histogram( aes(x = lfc1, y = ..count..), fill = "red")
p = p + geom_histogram( aes(x = lfc2, y = -..count..), fill = "green")
p = p + facet_grid(Gender+.~Study+Tissue, scales = 'free') + ylab('Gene count') + xlab('Log Fold Change')
p
```
### Overlap statistics
overlap between different brain regions in female
```{r ovlp.stats.up.female}
dexp.enrich.plots = dexp %>%
  dplyr::filter(Model %in% c('Diagnosis.Gender'), 
                adj.P.Val <= 0.05) %>%
  plyr::dlply(.(Gender, Direction), .fun = function(x){
    gr = x %>%
      plyr::dlply(.(Tissue), .fun = function(x){unique(x$ensembl_gene_id)}) %>%
      plyr::ldply(.,function(x, gs, backgroundGenes){
        plyr::ldply(gs, .fun = function(y,x,backgroundGenes){
          CovariateAnalysis::fisherEnrichment(x,y,backgroundGenes)
        }, x, backgroundGenes) %>%
          dplyr::rename(To = Tissue)
      }, ., background.genes) %>%
      dplyr::rename(From = Tissue) %>%
      dplyr::mutate(FDR = p.adjust(pval, method = 'BH')) %>%
      dplyr::select(From, To, Odds.Ratio, FDR, noverlap) %>%
      igraph::graph_from_data_frame()

    or.mat = igraph::as_adj(gr, type = 'both', attr = 'Odds.Ratio') %>% as.matrix()
    or.mat[is.infinite(or.mat)] = 0
    no.mat = igraph::as_adj(gr, type = 'both', attr = 'noverlap')

    if (unique(x$Direction) == 'Up'){
      col = c('white','red')
    } else {
      col = c('white','darkgreen')
    }
    ComplexHeatmap::Heatmap(as.matrix(or.mat), 
                            col= circlize::colorRamp2(c(0,40),col),
                            cell_fun = function(j, i, x, y, widht, height, fill){
                              no.mat[i,j] %>%
                                format(digit = 1) %>%
                                as.character() %>%
                                grid.text(x, y)
                            },
                            name = 'Odds Ratio',
                            column_title = paste(unique(x$Direction), 
                                                 'regulated in', 
                                                 unique(x$Gender)))
  })
purrr::walk(dexp.enrich.plots, function(x){ print(x) })
```
### Store in synapse
```{r syn.store}
# Get github commit
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Network", ref="branch", refName='AMPAD_manuscript')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('manuscriptFigures/',thisFileName))

obj = rmarkdown::render('diffExpFigure.Rmd', output_format = 'html_document')
obj = File(obj, name = 'Differential expression summary', parentId = 'syn11703045')
obj = synStore(obj, executed = thisFile)
# wikiPg = WikiPage(owner = obj, title = 'Figure1')
wikiPg = synGetWiki(obj)
wikiPg@properties$markdown = paste0("${preview?entityId=",obj$properties$id,"}")
wikiPg = synStore(wikiPg)
```