### 2018-01-27 code/Rmd/DEXP_ENRICH.Rmd
Added markdown to perform enrichment analysis for differentially expressed genes

---
title: "Enrichment analysis for differential expression results"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./DEXP_ENRICH.Rmd",
                                 parentId = "syn8672415",
                                 entityName = 'Enrichment analysis of all differentialy expressed genes')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(biomaRt)

library(synapseClient)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synapseLogin()

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn8672415';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of reprocessed rnaseq from all 7 brain regions';

thisFileName <- 'DEXP_ENRICH.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_reprocessing')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

## Data download
### Get summary differential expression results from synapse
```{r get.diffExp}
dexp = downloadFile('syn11180450')
all.used.ids = c(all.used.ids, 'syn11180450')

dexp$Direction = 'NONE'
dexp$Direction[dexp$adj.P.Val <= 0.05 & dexp$logFC >= log2(1.2)] = 'UP'
dexp$Direction[dexp$adj.P.Val <= 0.05 & dexp$logFC <= -log2(1.2)] = 'DOWN'
dexp$Tissue[dexp$Tissue == 'CER'] = 'CBE'

# Get hgnc symbol from biomart
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
                host = "www.ensembl.org", 
                dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", 
                       values = unique(dexp$ensembl_gene_id),
                       mart = mart)
dexp = dexp %>%
  dplyr::select(-hgnc_symbol) %>%
  dplyr::left_join(Ensemble2HGNC)

bck.genes = unique(dexp$hgnc_symbol)
```

### Get all gene sets
```{r get.gs}
# Get enrichr gene sets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c('syn4867851')
GeneSets = GeneSets[c('BioCarta', 'Chromosome_Location', 'Cross_Species_Phenotype', 
                      'Disease_Signatures_from_GEO_down', 'Disease_Signatures_from_GEO_up',
                      'ENCODE_TF_ChIP-seq', 'GO_Biological_Process', 
                      'KEGG_2015', 'Mouse_Gene_Atlas', 'OMIM_Disease', 'OMIM_Expanded', 'Panther', 
                      'Reactome', 'TargetScan_microRNA', 'TF-LOF_Expression_from_GEO', 'TRANSFAC_and_JASPAR_PWMs',
                      'WikiPathways', 'Single_Gene_Perturbations_from_GEO_up',
                      'Single_Gene_Perturbations_from_GEO_down', 'Aging_Perturbations_from_GEO_down',
                      'Aging_Perturbations_from_GEO_up', 'dbGaP', 'Disease_Perturbations_from_GEO_down',
                      'Disease_Perturbations_from_GEO_up', 'DrugMatrix', 
                      'Drug_Perturbations_from_GEO_down', 'Drug_Perturbations_from_GEO_up', 
                      'GTEx_Tissue_Sample_Gene_Expression_Profiles_down',
                      'GTEx_Tissue_Sample_Gene_Expression_Profiles_up', 'Jensen_DISEASES', 
                      'RNA-Seq_Disease_Gene_and_Drug_Signatures_from_GEO')]
gs = list()
gs$HypothesisFree = do.call(c, GeneSets) %>%
  lapply(function(y){
    str_split(y, '\\,') %>% 
      sapply(function(x){ x[[1]] })
  })

tmp = synTableQuery('select * from syn11707052')@values
all.used.ids = c(all.used.ids, 'syn11707052')

gs$CellType = tmp %>%
  dplyr::filter(GeneSet %in% grep('Zhang:', unique(tmp$GeneSet), value = T)) %>%
  plyr::dlply(.(GeneSet), .fun = function(x){
    unique(x$external_gene)
  })

gs$HypothesisDriven = tmp %>%
  dplyr::filter(!(GeneSet %in% grep('Zhang:', unique(tmp$GeneSet), value = T))) %>%
  plyr::dlply(.(GeneSet), .fun = function(x){
    unique(x$external_gene)
  })

gs = purrr::map(gs, function(geneSet, bckGenes){
  geneSet = purrr::map(geneSet, function(x, bckGenes){
    intersect(x, bckGenes) %>% unique()
  }, bckGenes)

  n = sapply(geneSet, length)
  geneSet = geneSet[n >= 10 & n <= 3000]
}, bck.genes)
```

### Perform enrichment analysis
```{r enrich.de}
enrich.results = dexp %>%
  dplyr::filter(Direction != 'NONE') %>%
  plyr::ddply(.(Model, Tissue, Comparison, Gender, Direction), .fun = function(x, gs, bckGenes){
    gsToTest = setdiff(unique(x$hgnc_symbol),'')
    enrich = plyr::ldply(gs, .fun = function(y, gsToTest, bckGenes){
      plyr::ldply(y, .fun = function(gsToTest1, gsToTest, bckGenes){
        CovariateAnalysis::fisherEnrichment(gsToTest1, gsToTest, bckGenes)
      }, gsToTest, bckGenes, .id = 'SetName', .parallel = T, .paropts = list(.packages = c('CovariateAnalysis')))
    }, gsToTest, bckGenes, .id = 'Category')
  }, gs, bck.genes)

fwrite(enrich.results, file = 'enrichmentResults.tsv', sep = '\t')
obj = File('enrichmentResults.tsv', name = 'Enrichment Results (for all differential expression)', parentId = parentId)
obj = synStore(obj, executed = thisFile, used = all.used.ids, 
               activityName = activityName, activityDescription = activityDescription)
```